---
title: "Cleaning_PreProcessing"
author: "Luis Jaime Vidal Jordi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This RMarkdown contains all the code elaborated for the Luis' Master Thesis. To follow the code it is highly advisable to previously read the thesis.

```{r}
# Some Libraries
library(INLA)
library(tidyverse)
library(magrittr)
library(xtable)
library(cowplot)
library(GGally)
```


## Importing Database

In this section we load the required data.

The Extreme European Climate Index data has been extracted from [E3CI WebPage](https://e3ci.dataclime.com/).


#### E3CI - Temporary data

We can load the data for one year, but the main idea is to work with the country data (we are going to fit spatial and spatio-temporal models):

```{r}
jan_1990 <- read.csv("Tiempo/1990/era5_country_january1990.csv")
feb_1990 <- read.csv("Tiempo/1990/era5_country_february1990.csv")
march_1990 <- read.csv("Tiempo/1990/era5_country_march1990.csv")
april_1990 <- read.csv("Tiempo/1990/era5_country_april1990.csv")
may_1990 <- read.csv("Tiempo/1990/era5_country_may1990.csv")
jun_1990 <- read.csv("Tiempo/1990/era5_country_june1990.csv")
jul_1990 <- read.csv("Tiempo/1990/era5_country_july1990.csv")
aug_1990 <- read.csv("Tiempo/1990/era5_country_august1990.csv")
sept_1990 <- read.csv("Tiempo/1990/era5_country_september1990.csv")
oct_1990 <- read.csv("Tiempo/1990/era5_country_october1990.csv")
nov_1990 <- read.csv("Tiempo/1990/era5_country_november1990.csv")
dec_1990 <- read.csv("Tiempo/1990/era5_country_december1990.csv")

year_1990 <- rbind(jan_1990,feb_1990,march_1990,april_1990,may_1990,jun_1990,jul_1990,aug_1990,sept_1990,oct_1990,nov_1990,dec_1990)

# Esto sería lo que queremos para trabajar:
year_1990_sum <- year_1990 %>% group_by(., Country) %>% summarise(.,"E3CI" = mean(E3CI))

# Luego faltaría añadir las variables de interés y filtrar los países que no queremos.
rm(list = ls())
```

#### E3CI - Country Data

If we want to use the information by years, maybe it is more convenient to load all the data by countries and so we have all the years available that we want. Let's clean Spain and if it works we repeat it for the rest:

```{r}
spain <- read_csv("Paises/era5_country_spain.csv")
```

Test whether what I want to do works for Spain:

```{r}
spain_sum <- spain %>% separate(Month, into = c("Month", "Year"), sep = "-") %>% group_by(Year) %>% summarise("E3CI" = mean(E3CI))
```

I see that it is, so I load all the data:

```{r, message=FALSE}
albania <- read_csv("Paises/era5_country_albania.csv")
andorra <- read_csv("Paises/era5_country_andorra.csv")
austria <- read_csv("Paises/era5_country_austria.csv")
belarus <- read_csv("Paises/era5_country_belarus.csv")
belgium <- read_csv("Paises/era5_country_belgium.csv")
bosnia_herz <- read_csv("Paises/era5_country_bosnia_and_herzegovina.csv")
bulgaria <- read_csv("Paises/era5_country_bulgaria.csv")
croatia <- read_csv("Paises/era5_country_croatia.csv")
cyprus <- read_csv("Paises/era5_country_cyprus.csv")
czechia <- read_csv("Paises/era5_country_czechia.csv")
denmark <- read_csv("Paises/era5_country_denmark.csv")
estonia <- read_csv("Paises/era5_country_estonia.csv")
finland <- read_csv("Paises/era5_country_finland.csv")
france <- read_csv("Paises/era5_country_france.csv")
germany <- read_csv("Paises/era5_country_germany.csv")
greece <- read_csv("Paises/era5_country_greece.csv")
hungary <- read_csv("Paises/era5_country_hungary.csv")
iceland <- read_csv("Paises/era5_country_iceland.csv")
ireland <- read_csv("Paises/era5_country_ireland.csv")
italy <- read_csv("Paises/era5_country_italy.csv")
latvia <- read_csv("Paises/era5_country_latvia.csv")
lebanon <- read_csv("Paises/era5_country_lebanon.csv")
lithuania <- read_csv("Paises/era5_country_lithuania.csv")
luxembourg <- read_csv("Paises/era5_country_luxembourg.csv")
malta <- read_csv("Paises/era5_country_malta.csv")
moldova <- read_csv("Paises/era5_country_moldova.csv")
montenegro <- read_csv("Paises/era5_country_montenegro.csv")
morocco <- read_csv("Paises/era5_country_morocco.csv")
netherlands <- read_csv("Paises/era5_country_netherlands.csv")
north_maced <- read_csv("Paises/era5_country_north_macedonia.csv")
norway <- read_csv("Paises/era5_country_norway.csv")
poland <- read_csv("Paises/era5_country_poland.csv")
portugal <- read_csv("Paises/era5_country_portugal.csv")
romania <- read_csv("Paises/era5_country_romania.csv")
serbia <- read_csv("Paises/era5_country_serbia.csv")
slovakia <- read_csv("Paises/era5_country_slovakia.csv")
slovenia <- read_csv("Paises/era5_country_slovenia.csv")
spain <- read_csv("Paises/era5_country_spain.csv")
sweden <- read_csv("Paises/era5_country_sweden.csv")
switzerland <- read_csv("Paises/era5_country_switzerland.csv")
tunisia <- read_csv("Paises/era5_country_tunisia.csv")
ukraine <- read_csv("Paises/era5_country_ukraine.csv")
united_kingdom <- read_csv("Paises/era5_country_united_kingdom.csv")
```

I am going to put them in a single object (list) called *countries*.

```{r, message=FALSE}
countries <- list(
  Albania = albania,
  Andorra = andorra,
  Austria = austria,
  Belarus = belarus,
  Belgium = belgium,
  `Bosnia and Herzegovina` = bosnia_herz,
  Bulgaria = bulgaria,
  Croatia = croatia,
  Cyprus = cyprus,
  Czechia = czechia,
  Denmark = denmark,
  Estonia = estonia,
  Finland = finland,
  France = france,
  Germany = germany,
  Greece = greece,
  Hungary = hungary,
  Iceland = iceland,
  Ireland = ireland,
  Italy = italy,
  Latvia = latvia,
  Lebanon = lebanon,
  Lithuania = lithuania,
  Luxembourg = luxembourg,
  Malta = malta,
  Moldova = moldova,
  Montenegro = montenegro,
  Morocco = morocco,
  Netherlands = netherlands,
  `North Macedonia` = north_maced,
  Norway = norway,
  Poland = poland,
  Portugal = portugal,
  Romania = romania,
  Serbia = serbia,
  Slovakia = slovakia,
  Slovenia = slovenia,
  Spain = spain,
  Sweden = sweden,
  Switzerland = switzerland,
  Tunisia = tunisia,
  Ukraine = ukraine,
  `United Kingdom` = united_kingdom
)

names(countries)
```

The data corresponding to each country contains more information than we want to use. In particular the variable corresponding to the time (Month) comes in *mm-yyyy* format.

Since the first efforts we will make will be focused on specific years (spatial model only) we will reduce the information to that: *index-year* for each country:

```{r, message=FALSE}
# Object to store the aggregate df per year
countries_summ <- list(
  Albania = NULL,
  Andorra = NULL,
  Austria = NULL,
  Belarus = NULL,
  Belgium = NULL,
  `Bosnia and Herzegovina`= NULL,
  Bulgaria = NULL,
  Croatia = NULL,
  Cyprus = NULL,
  Czechia = NULL,
  Denmark = NULL,
  Estonia = NULL,
  Finland = NULL,
  France = NULL,
  Germany = NULL,
  Greece = NULL,
  Hungary = NULL,
  Iceland = NULL,
  Ireland = NULL,
  Italy = NULL,
  Latvia = NULL,
  Lebanon = NULL,
  Lithuania = NULL,
  Luxembourg = NULL,
  Malta = NULL,
  Moldova = NULL,
  Montenegro = NULL,
  Morocco = NULL,
  Netherlands = NULL,
  `North Macedonia` = NULL,
  Norway = NULL,
  Poland = NULL,
  Portugal = NULL,
  Romania = NULL,
  Serbia = NULL,
  Slovakia = NULL,
  Slovenia = NULL,
  Spain = NULL,
  Sweden = NULL,
  Switzerland = NULL,
  Tunisia = NULL,
  Ukraine = NULL,
  `United Kingdom` = NULL
)
```

We are also going to add a column to each country to indicate precisely which country it is, so that we can distinguish between them when we combine everything in a single table:

```{r}
for(i in 1:length(countries)){
  countries_summ[[i]] <- countries[[i]]  %>% separate(Month, into = c("Month", "Year"), sep = "-") %>% group_by(Year) %>% summarise("Country" = names(countries)[i],"E3CI" = mean(E3CI)) %>% relocate(Country, .before = Year)}
```

Finally we can combine all countries in a single data frame:

```{r}
year_data <- NULL
for(i in 1:length(countries)){
  year_data <- rbind(year_data,countries_summ[[i]])
}
```


## Covariates

[Covariates Webpage](https://ourworldindata.org/co2-and-greenhouse-gas-emissions)

### CO2 per capita

We load the data

```{r,message=FALSE}
co2 = read_csv("Covariates/co-emissions-per-capita.csv")
str(co2)
summary(co2)
```

We see that we have years too many. In this case we only want from 1981 onwards. We have co2 until 2022 (not 2024).

First let's simplify the c02 data to what we need (years and countries).

```{r}
# We need to make sure that the country names in the data we have and the ones we want to add are consistent. I have coded above to match :)

# It must be the same number
intersect(names(countries_summ),co2$Entity) %>% length()
length(names(countries_summ))



# All that remains is to filter the years and countries of interest.
co2_filt <- co2 %>% filter(Year >= 1981 & (Entity %in% names(countries_summ)))

# We change the name of the variables
colnames(co2_filt) = c("Country","Code","Year","CO2_emissions")

# IMPORTANT BEFORE ADDING THE COLUMN: check that the rows match in year and country. Remember that we only have data up to 2022 of emissions so:
year_data_to2022 <- year_data[which((year_data$Year != 2023)&(year_data$Year != 2024)),]
length(co2_filt$Year) == length(year_data_to2022$Year) # Misma cantidad de observaciones
# Check rows order:
unique(co2_filt$Year == year_data_to2022$Year) # Años coincidentes
unique(co2_filt$Country == year_data_to2022$Country) # Países coincidentes

# We can now do the cbind (or add with mutate) in a safe way. We also add the code since:
year_data_to2022 <- cbind(year_data_to2022,co2_filt[,c("Code","CO2_emissions")])


# I rearrange some columns for the convenience of displaying the data.
year_data_to2022 <- year_data_to2022 %>% relocate(Code, .before = Country)
```

#### Pre-Processing CO2

First of all we check if there are missing values:

```{r}
sum(is.na(year_data_to2022$CO2_emissions))
```

As it is not the case, we proceed with the pre-processing.

These data are actually time series, so they should be pre-processed accordingly.

We convert it into a time series:
```{r}
co2_data <- year_data_to2022 %>% dplyr::select(Country, Year, CO2_emissions)
co2_data$Country <- as.factor(co2_data$Country)
str(co2_data)
```

```{r}
ggplot(co2_data, aes(x = Year, y = CO2_emissions, color = Country, group = Country)) +
  geom_line() +
  labs(x = "Año", y = "Emisiones de CO2", color = "País") +
  theme_minimal() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("Emisiones de CO2 por País a lo largo del Tiempo")

```

```{r}
# Explore some random countries
co2_spain <- ts(co2_data$CO2_emissions[which(co2_data$Country=="Spain")],
                start = 1981, end = 2022)
co2_morocco <- ts(co2_data$CO2_emissions[which(co2_data$Country=="Morocco")],
                start = 1981, end = 2022)
co2_italy <- ts(co2_data$CO2_emissions[which(co2_data$Country=="Italy")],
                start = 1981, end = 2022)
ts.plot(co2_spain)
ts.plot(co2_morocco)
ts.plot(co2_italy)

acf(co2_spain)
acf(co2_morocco)
```
In general, some trend, so that probably is advisible to take differences.

We can create a function to check for general seasonality or stationarity by country:
```{r}
library(tseries)
check_stationarity <- function(data, country) {
  ts_data <- ts(data[(data$Country == country), "CO2_emissions"], start = 1981, end = 2022)
  adf_test <- adf.test(ts_data)
  return(adf_test$p.value)
}

for(i in unique(co2_data$Country)){
  p_val <- check_stationarity(co2_data,i)
  print(paste0("p-value of ",i, " is ", p_val))
}

```

There are some with a value of 0.06. But none of them goes below the 0.05 significance level. Then, we can differentiate them all.

We would like to differentiate them in a way that I can add a column next to the original one (keeping order and so on) that allows me to have the two series in parallel.

```{r}
length(co2_spain)
co2_spain
diff(co2_spain)
```

We join to `co2_data` the vector containing the corresponding differenced series. We lose one year. 

```{r}
difference_series <- function(data, country, CO2_diff) {
  country_data <- data %>% filter(Country == country)
  CO2_diff <- c(CO2_diff,c(NA, diff(country_data$CO2_emissions)))
  return(CO2_diff)
}

CO2_diff <- NULL
for(i in unique(co2_data$Country)){
  CO2_diff <- difference_series(data = co2_data,country = i,CO2_diff = CO2_diff)
}

if(length(co2_data$CO2_emissions)==length(CO2_diff)){
  co2_data$CO2_diff <- CO2_diff 
}
```

```{r}
year_data_to2022$CO2_diff <- co2_data$CO2_diff
```

### GDP (Gross Domestic Product)

The following selected covariate is the gross domestic product of the countries. The webpage is the same as for the CO2 covariate. 

```{r, message=FALSE, warning=FALSE}
gdp = read_csv("Covariates/gdp-per-capita-maddison2.csv")
str(gdp)

sum(is.na(gdp$`GDP per capita`)) # No missing values en GDP
```

However, looking at the dataframe there is a slightly odd column that says *annotations*. 

```{r}
head(gdp$`900793-annotations`)
```

Check:

```{r}
sum(is.na(gdp$`900793-annotations`)) == nrow(gdp)
```

As they are all NAs we can get rid of it:

```{r}
gdp <- gdp[,1:4]
# We also change the column names
colnames(gdp) = c("Country","Code","Year","GDP")
```

The above process would now have to be repeated:

```{r}
# We need to match our country labels to the gdp data so that we can ‘cbind’ later. We have 40 out of 41 matches.

intersect(names(countries_summ),gdp$Country) %>% length()
length(names(countries_summ))
```

However we see that there is one country for which we do not have the GDP. There is no GDP data for **Andorra**. I suppose we could look it up separately or eliminate the country from the study.

For the moment I will artificially create the variable with the years and we will give a value of NA to its GDP. Later on, if data is no available we will eliminate it.

```{r}
andorra_gdp <- data.frame(Country = rep("Andorra",length(1950:2022)),Code = rep("AND",length(1950:2022)), Year = 1950:2022, GDP = rep(NA,length(1950:2022)))
```

It is important to put it in the same position as CO2 to avoid problems later (after Albania).

Albania is the second **gdp** country. Therefore, it is important to place it in the same position as CO2 to avoid problems later on (after Albania).

```{r}
# Primer pais que no es albania en posicion:
indice1 = sum(gdp$Country == "Afghanistan" | gdp$Country == "Albania") + 1
indice2 = indice1 + 1 
gdp <- rbind(gdp[1:indice1,],andorra_gdp,gdp[indice2:nrow(gdp),])
```

Let's see if we now have the inter-section of fully matching labels (41):

```{r}
intersect(names(countries_summ),gdp$Country) %>% length()
```

Now that we are the same, the only thing left to do is to filter the years and countries of interest:
```{r}
gdp_filt <- gdp %>% filter(Year >= 1981 & (Country %in% names(countries_summ)))
```

New problem. There is only data up to 2018. So we add NA to all the following years in case we get that information later.

Before doing so, it is important to realize that in this case we have a country with fewer observations than we need:

```{r}
table(gdp_filt$Country)
```

Slovakia has 34 rows. Let's see which ones are missing:

```{r}
cond = (gdp_filt$Country == "Slovakia")
gdp_filt[cond,"Year"]
```

We see that the first 4 years are missing. We add them by hand

```{r}
new_rows_slovakia <- data.frame(
  Country = rep("Slovakia",4),
  Code = rep("SVK",4),
  Year = 1981:1984,
  GDP = rep(NA,4)
)

# we need the position in which we want to add the new values
pos <- which((gdp_filt$Year == 1985)&(gdp_filt$Country == "Slovakia"))
sup <- gdp_filt[1:(pos-1),]
inf <- gdp_filt[pos:(nrow(gdp_filt)),]
gdp_filt <- rbind(sup, new_rows_slovakia, inf)
```

```{r}
# All well and in order
table(gdp_filt$Country)
```


```{r}
# IMPORTANT BEFORE ADDING THE COLUMN: check that the rows match in year and country.
# Check the order of the rows:
unique(gdp_filt$Year == year_data_to2022$Year) # Coincident years
unique(gdp_filt$Country == year_data_to2022$Country) # and countries

# We can now do the cbind (or add with mutate) safely:
clean_data <- cbind(year_data_to2022,gdp_filt[,"GDP"])
```

We already have some initial data to work with. Obviously there are NA, so we will have to obtain the data externally or impute them with some criteria.


It seems that there are some outliers, but we decide to keep it in the data:

```{r}
plot(density(year_data_to2022$E3CI))
boxplot(year_data_to2022$E3CI)
```

```{r}
quantile(year_data_to2022$E3CI)
```


#### Pre-Processing - Seasonality
Again, first let us check the missing values. 

```{r}
sum(is.na(clean_data$GDP))
```

The Andorra missing values will be keep it just in case we get the data later on. However, we have 4 more NA which correspond to Slovakia.

```{r}
Slov_NA <- which(clean_data$Country == "Slovakia")
clean_data$GDP[Slov_NA]
```

We impute them using splines through the library `imputeTS`, which is a library that allows easy imputation of missing values in time series.

```{r}
library(imputeTS)
```

The function we use is `na_interpolation(x, option = "spline")`:

```{r}
imputed_slov <- imputeTS::na_interpolation(clean_data$GDP[Slov_NA], option = "spline") %>% print()
```
Then,
```{r}
clean_data$GDP[Slov_NA] <- imputed_slov
```

Check:

```{r}
# The only missing values now should be the corresponding to Andorra. 
Ando <- which(clean_data$Code == "AND")
sum(is.na(clean_data$GDP[-Ando]))
```



Now, we want to treat it as a time series
```{r}
gdp_data <- clean_data %>% dplyr::select(Country, Year, GDP)
gdp_data$Country <- as.factor(gdp_data$Country)
str(gdp_data)
```

```{r}
ggplot(gdp_data, aes(x = Year, y = GDP, color = Country, group = Country)) +
  geom_line() +
  labs(x = "Year", y = "GDP", color = "Country") +
  theme_minimal() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("DGP (per capita) by country along time")

```

```{r}
# Algunos países randoms
gdp_spain <- ts(gdp_data$GDP[which(gdp_data$Country=="Spain")],
                start = 1981, end = 2022)
gdp_morocco <- ts(gdp_data$GDP[which(gdp_data$Country=="Morocco")],
                start = 1981, end = 2022)
gdp_italy <- ts(gdp_data$GDP[which(gdp_data$Country=="Italy")],
                start = 1981, end = 2022)
ts.plot(gdp_spain)
ts.plot(gdp_morocco)
ts.plot(gdp_italy)

acf(na.omit(gdp_spain))
acf(na.omit(gdp_morocco))
```
We can create a function to check for general seasonality or stationarity by country.
```{r}
library(tseries)
check_stationarity_gdp <- function(data, country) {
  ts_data <- ts(na.omit(data[(data$Country == country), "GDP"]), start = 1981, end = 2018)
  adf_test <- adf.test(ts_data)
  return(adf_test$p.value)
}

for(i in unique(gdp_data$Country)){
  if(i == "Andorra"){
    print("De momento no tenemos GDP de Andorra")
    next
  }
  p_val <- check_stationarity_gdp(gdp_data,i)
  print(paste0("p-value of ",i, " is ", p_val))
  
}
```

In this case only one country is close to 0.05 (0.08.). We differentiate them all.

I would like to differentiate them in a way that I can add a column next to the original one (keeping order and so on) that allows me to have the two series in parallel.

We have already done it in CO2 and we have seen that we lose the first value. In this case a NA is not generated (but there is one value less. We will assume that it is the first one that is lost). Moreover, we have to take into account that we have 4 NA at the end. But it seems that they are left to differentiate so there will be no problems.

```{r}
length(gdp_spain)
gdp_spain
diff((gdp_spain))
```
We join to `gdp_data` the vector containing the corresponding differenced series. We lose one year. 

```{r}
difference_series_gdp <- function(data, country, gdp_diff) {
  country_data <- data %>% filter(Country == country)
  vec <- c(gdp_diff,c(NA, diff(country_data$GDP)))
  return(vec)
}

gdp_diff <- NULL
for(i in unique(gdp_data$Country)){
  if(i == "Andorra"){
    gdp_diff <- c(gdp_diff, rep(NA, length(which(gdp_data$Country == i))))
    next
  }
  gdp_diff <- difference_series_gdp(data = gdp_data,country = i,gdp_diff = gdp_diff)
}
```

Then if everything is correct we add the fixed column.
```{r}
if(length(gdp_data$GDP)==length(gdp_diff)){
  gdp_data$GDP_diff <- gdp_diff 
}

clean_data$GDP_diff <- gdp_data$GDP_diff
```


### GHG - Greenhouse gas esmissions

The same procedure is repeated for the Greenhouse Emissions.

```{r}
ghg <- read_csv("Covariates/per-capita-ghg-emissions.csv")
```

```{r}
# We need to make sure that the country names in the data we have and those we want to add are consistent. I have coded above to match

# We should get the same number
intersect(names(countries_summ),ghg$Entity) %>% length()
length(names(countries_summ))

# All that remains is to filter the years and countries of interest.
ghg_filt <- ghg %>% filter(Year >= 1981 & (Entity %in% names(countries_summ)))

# We change the name of the variables
colnames(ghg_filt) = c("Country","Code","Year","GHG")

# IMPORTANT BEFORE ADDING THE COLUMN: check that the rows match in year and country. Remember that we only have data up to 2022 of emissions so:
length(ghg_filt$Year) == length(clean_data$Year) # Same number of observations
# Check rows order:
unique(ghg_filt$Year == clean_data$Year) # Coincident Years
unique(ghg_filt$Country == clean_data$Country) # Coincident Countries

# We can now do the cbind (or add with mutate) in a safe way. We also add the code since:
clean_data <- cbind(clean_data,ghg_filt[,"GHG"])
```

We have already placed the variable in `clean_data`. Now it is time to see if it needs to be differentiated.

#### Pre-Processing GHG

```{r}
ghg_data <- clean_data %>% dplyr::select(Country, Year, GHG)
ghg_data$Country <- as.factor(ghg_data$Country)
str(ghg_data)

# We check that we have no NA (we have already checked before that all the desired countries were present).
unique(is.na(ghg_data)) == c(FALSE,FALSE,FALSE)
```

```{r}
ggplot(ghg_data, aes(x = Year, y = GHG, color = Country, group = Country)) +
  geom_line() +
  labs(x = "Year", y = "GHG", color = "Country") +
  theme_minimal() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("GHG (per capita) by country along time")
```

```{r}
# Some random countries
ghg_spain <- ts(ghg_data$GHG[which(ghg_data$Country=="Spain")],
                start = 1981, end = 2022)
ghg_morocco <- ts(ghg_data$GHG[which(ghg_data$Country=="Morocco")],
                start = 1981, end = 2022)
ghg_italy <- ts(ghg_data$GHG[which(ghg_data$Country=="Italy")],
                start = 1981, end = 2022)
ts.plot(ghg_spain)
ts.plot(ghg_morocco)
ts.plot(ghg_italy)

acf(ghg_spain)
acf(ghg_morocco)
```
We can create a function to check the general seasonality or stationarity by country also in ghg.
```{r}
library(tseries)
check_stationarity_ghg <- function(data, country) {
  ts_data <- ts(data[(data$Country == country), "GHG"], start = 1981, end = 2022)
  adf_test <- adf.test(ts_data)
  return(adf_test$p.value)
}

for(i in unique(ghg_data$Country)){
  p_val <- check_stationarity_ghg(ghg_data,i)
  print(paste0("p-value of ",i, " is ", p_val))
  
}
```

In this case only one country is close to 0.05 (0.06). We differentiate them all.

I would like to differentiate them in a way that I can add a column next to the original one (keeping order and so on) that allows me to have the two series in parallel.

We have already done it in CO2 and we have seen that we lose the first value. In this case a NA is not generated (but there is one value less. We will assume that it is the first one that is lost). Moreover, we have to take into account that we have 4 NA at the end. But it seems that they are left to differentiate so there will be no problems.

```{r}
length(ghg_spain)
ghg_spain
diff((ghg_spain))
length(diff((ghg_spain)))
```

We join to `gdp_data` the vector containing the corresponding differenced series. We lose one year. 

```{r}
difference_series_ghg <- function(data, country, ghg_diff) {
  country_data <- data %>% filter(Country == country)
  vec <- c(ghg_diff,c(NA, diff(country_data$GHG)))
  return(vec)
}

ghg_diff <- NULL
for(i in unique(ghg_data$Country)){
  ghg_diff <- difference_series_ghg(data = ghg_data,country = i,ghg_diff = ghg_diff)
}
```

```{r}
if(length(ghg_data$GHG)==length(ghg_diff)){
  print("All ok")
  ghg_data$GHG_diff <- ghg_diff 
}

clean_data$GHG_diff <- ghg_data$GHG_diff
```

Finally we convert to numeric the col year:

```{r}
clean_data$Year <- as.numeric(clean_data$Year)
```

### FA - Forest area

The Land Area has been extracted from other website:

[Forest Area Website](https://www.globalforestwatch.org/blog/data/forest-extent-2020-gfw-umd-forest-resources-assessment/)

Forest Area:
```{r}
land_area <- read.csv2("Covariates/Modified_Data_Land_Area.csv", sep = ";", header = TRUE, dec = ".")
str(land_area)
```

- Modify year column name:
```{r}
land_area <- land_area %>%
  rename_with(~gsub("^X", "", .), starts_with("X"))
```

- We can remove the columns that are not of interest and relabel the other ones
```{r}
land_area <- land_area %>% dplyr::select(-c("Indicator.Name","Indicator.Code", "Country.Name.1","Indicator.Name.1","Country.Code.1")) %>% 
  rename(Country = Country.Name, Code = Country.Code) 
str(land_area)
```

- Remove the years that are out of the scope (all NAs from 1960 to 1989)
```{r}
land_area <- land_area[,-c(3:23)]
str(land_area)
```

Now the data must be converted to long format to make it a time series like the others:
```{r}
land_area <- land_area %>% pivot_longer(cols = -c(Country,Code),
                                        names_to = "Year",
                                        values_to = "Forest_area")

land_area <- as.data.frame(land_area)
```

- And now we stay with the countries that interest us:
```{r}
land_area <- land_area %>% filter(Code %in% unique(clean_data$Code))
# Comprobamos que tenemos la misma cantidad de paises:
length(unique(clean_data$Code)) == length(unique(land_area$Code))
# Ojo cuidado que no estan en el mismo orden

land_area$Year <- as.numeric(land_area$Year)
```

Now let's add the column to `clean data` in a coherent and ordered way.

- First we add 2022 (I put it the same as 2021).
```{r}
# First we add NAs for the year 2022
n = nrow(land_area)
j = 0
for(i in 1:(n)){
  if(land_area[i+j,"Year"] == 2021){
    add2022 <- data.frame(Country = land_area[i+j,"Country"],
                      Code = land_area[(i+j),"Code"],
                      Year = land_area[(i+j),"Year"] + 1,
                      Forest_area = NA)
    inf <- land_area[1:(i+j),]
    sup <- land_area[(i+j+1):nrow(land_area),]
    land_area <- rbind(inf,add2022,sup)
    j = nrow(land_area) - n
  }
}
# Two extra rows are added which we can (should) remove:

land_area <- land_area[1:(nrow(land_area)-2),]

nrow(clean_data) == nrow(land_area)
```


#### Pre-Processing FA
We have a lot of missing values. As with GDP we impute the missing values with interpolation.

```{r}
for(i in unique(land_area$Country)){
  cond <- which(land_area$Country == i)
  to_impute <- land_area[cond,"Forest_area"]
  imputed <- na_interpolation(to_impute,option = "spline")
  land_area[cond,"Forest_area"] <- imputed
}
```

Check:
```{r}
sum(is.na(land_area))
```

- Now we must consistently add the column to clean data (matching year and country). 
```{r}
Forest_Area <- c()
for(i in 1:nrow(clean_data)){
  condition <- which((land_area$Code == clean_data[i,"Code"])&(land_area$Year == clean_data[i,"Year"]))
  Forest_Area <- c(Forest_Area, land_area[condition,"Forest_area"])
}
Forest_Area <- as.numeric(Forest_Area)

clean_data <- cbind(clean_data,Forest_Area)
```

```{r}
ggplot(clean_data, aes(x = Year, y = Forest_Area, color = Country, group = Country)) +
  geom_line() +
  labs(x = "Year", y = "GDP", color = "Country") +
  theme_minimal() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("DGP (per capita) by country along time")

```


**Stationary process?**

```{r}
# Some Random Countries
Fa_spain <- ts(land_area$Forest_area[which(land_area$Country=="Spain")],
                start = 1981, end = 2022)
Fa_morocco <- ts(land_area$Forest_area[which(land_area$Country=="Morocco")],
                start = 1981, end = 2022)
Fa_italy <- ts(land_area$Forest_area[which(land_area$Country=="Italy")],
                start = 1981, end = 2022)
ts.plot(Fa_spain)
ts.plot(Fa_morocco)
ts.plot(Fa_italy)

acf(Fa_spain)
acf(Fa_morocco)
```

```{r}
library(tseries)
check_stationarity_fa <- function(data, country) {
  ts_data <- ts(data[(data$Country == country), "Forest_area"], start = 1981, end = 2022)
  adf_test <- adf.test(ts_data)
  return(adf_test$p.value)
}

for(i in unique(land_area$Country)){
  p_val <- check_stationarity_fa(land_area,i)
  print(paste0("p-value of ",i, " is ", p_val))
}
```


```{r}
length(Fa_spain)
Fa_spain
diff((Fa_spain))
length(diff((Fa_spain)))
```

```{r}
difference_series_fa <- function(data, country, FA_diff){
  country_data <- data %>% filter(Country == country)
  vec <- c(FA_diff,c(NA, diff(country_data$Forest_area)))
  return(vec)
}

FA_diff <- NULL
for(i in unique(land_area$Country)){
  FA_diff <- difference_series_fa(data = land_area,country = i,FA_diff = FA_diff)
}
length(FA_diff)
```

Then if everything is correct we add the fixed column
```{r}
if(length(land_area$Forest_area)==length(FA_diff)){
  print("All ok")
  land_area$FA_diff <- FA_diff 
}

clean_data$FA_diff <- land_area$FA_diff
```


### E3CI: Pre-processing

```{r}
ggplot(clean_data, aes(x = Year, y = E3CI, color = Country, group = Country)) +
  geom_line() +
  labs(x = "Year", y = "GDP", color = "Country") +
  theme_minimal() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("E3CI by country over time")

```

```{r}
library(tseries)
check_stationarity_E3CI <- function(data, country) {
  ts_data <- ts(data[(data$Country == country), "E3CI"], start = 1981, end = 2022)
  adf_test <- adf.test(ts_data)
  return(adf_test$p.value)
}

for(i in unique(clean_data$Country)){
  p_val <- check_stationarity_E3CI(clean_data,i)
  print(paste0("p-value of ",i, " is ", p_val))
}
```


```{r}
difference_series_E3CI <- function(data, country, E3CI_diff){
  country_data <- data %>% filter(Country == country)
  vec <- c(E3CI_diff,c(NA, diff(country_data$E3CI)))
  return(vec)
}

E3CI_diff <- NULL
for(i in unique(clean_data$Country)){
  E3CI_diff <- difference_series_E3CI(data = clean_data,country = i,E3CI_diff = E3CI_diff)
}
length(E3CI_diff)
```


```{r}
if(length(clean_data$E3CI)==length(E3CI_diff)){
  print("All ok")
  clean_data$E3CI_diff <- E3CI_diff 
}
```

### Missing values

So far we have more or less pre-processed the data we are going to work with. However, there are still some missing values. These are Andorra and the first year for each country, which were lost due to differentiation.

```{r}
sum(is.na(clean_data))
```


Up to this point we have not got the Andorra GDP, so we will remove it of the study.

```{r}
And <- which(clean_data$Country == "Andorra")
clean_data <- clean_data[-And,]
```

Then, now:
```{r}
sum(is.na(clean_data))
```

As we have differentiated our series, we have lost one year, but still in the data with NA. Let us remove it
```{r}
year_1981 <- which(clean_data$Year == 1981)
clean_data <- clean_data[-year_1981,]
```

Finally,

```{r}
sum(is.na(clean_data))
```


### Standardize all covariates

So far, we have all the covariates included in `clean_data` (differentiated and undifferentiated).

```{r}
str(clean_data)
```

Let us standardize the covariates to make comparable the coefficients in posterior sections, where INLA models are fitted.

```{r}
clean_data_nostand <- clean_data

cols_to_standardize <- c("CO2_emissions","CO2_diff","GDP","GDP_diff","GHG","GHG_diff",      "Forest_Area","FA_diff")

clean_data <- clean_data %>% 
  mutate(across(all_of(cols_to_standardize), ~ scale(.)))

colnames(clean_data) <- colnames(clean_data_nostand)
```



## Map

First of all we load some libraries

```{r, message=FALSE}
library(mapdata)
library(ggplot2)
library(maps)
library(sf)
library(sp)
```

We need to load a geometric object that contains the spatial information of our model, the **shapefiles**. The shapefile format is capable of storing data in the form of points, lines, or (multi)polygons. Shapefiles are employed to represent administrative boundaries, such as country, state, or county borders, as well as roads, rivers, lakes, and other geographical features.


In our case, we want to work with an arbitrary selection of countries of interest (mainly Europe).

A world shapefile can be obtained in the following website:

[World_shapefile](https://public.opendatasoft.com/explore/dataset/world-administrative-boundaries/export/)

```{r}
shp_world <- read_sf("world-administrative-boundaries/world-administrative-boundaries.shp")
```

In this data, the geometry structure is *Multilpolygon*

```{r}
st_geometry(shp_world)
```

Let us filter the shapefile to hold only the countries for which we have available climate data.

```{r}
head(clean_data)
```

Then

```{r}
shp_world_filtered <- shp_world %>% filter(color_code %in% unique(clean_data$Code))
```


However, the original map contains several instances (regions) labelled as the same countries.

```{r}
shp_world_filtered$color_code
length(shp_world_filtered$color_code)
length(unique(shp_world_filtered$color_code))
```

They are islands or other territories that are not in the main country region. As we want to fit spatio-temporal models, we will work only with the main country region.

-   GBR: 12 veces -\> 11 extra
-   FRA: 10 veces -\> 9 extra
-   DNK: 3 veces -\> 2 extra
-   PRT: 3 veces -\> 2 extra
-   NOR: 3 veces -\> 2 extra
-   NLD: 2 veces -\> 1 extra

We have 27 regions to spare. 

```{r}
cond_GBR <- which((shp_world_filtered$color_code == "GBR")&(shp_world_filtered$status != "Member State"))
cond_FRA <- which((shp_world_filtered$color_code == "FRA")&(shp_world_filtered$status != "Member State"))
cond_DNK <- which((shp_world_filtered$color_code == "DNK")&(shp_world_filtered$status != "Member State"))
cond_PRT <- which((shp_world_filtered$color_code == "PRT")&(shp_world_filtered$status != "Member State"))
cond_NOR <- which((shp_world_filtered$color_code == "NOR")&(shp_world_filtered$status != "Member State"))
cond_NLD <- which((shp_world_filtered$color_code == "NLD")&(shp_world_filtered$status != "Member State"))


shp_world_filtered_unique <- shp_world_filtered[-c(cond_DNK,cond_FRA,cond_GBR,cond_NLD,cond_NOR,cond_PRT),]
```

The `name` column of the `shp_world_filtered_unique` does not match with the `Country` variables of the covariates (`clean_data`). Let us fix this to avoid problems in the future works:

```{r}
for(i in unique(clean_data$Code)){
  cond_name <- which(shp_world_filtered_unique$color_code == i)
  cond_code <- which(clean_data$Code == i)
  corrected_name <- unique(clean_data$Country[cond_code])
  shp_world_filtered_unique$name[cond_name] = corrected_name
}
```


Then, a first visualization of our map:

```{r}
map1 <- ggplot(data = shp_world_filtered_unique) +
  geom_sf()

map1
```


The following function takes a year as input and filters from `shp_world_filtered_unique` the corresponding year data.

```{r}
#_diff because there is a version selecting the time series but without differenciate
select_year_diff <- function(year = 2000){
  Year <- NULL
  E3CI_diff <- NULL
  E3CI <- NULL
  CO2 <- NULL
  GDP <- NULL
  GHG <- NULL
  FA <- NULL

  for(i in shp_world_filtered_unique$color_code){
    cond <- which((clean_data$Code == i)&(clean_data$Year == year))
    E3CI_diff <- c(E3CI_diff,clean_data[cond,"E3CI_diff"])
    E3CI <- c(E3CI,clean_data[cond,"E3CI"])
    CO2 <- c(CO2, clean_data[cond,"CO2_diff"])
    GDP <- c(GDP, clean_data[cond,"GDP_diff"])
    GHG <- c(GHG, clean_data[cond,"GHG_diff"])
    FA <- c(FA, clean_data[cond,"FA_diff"])
    Year <- c(Year,year)}

  return(cbind(shp_world_filtered_unique,data.frame(E3CI_diff = E3CI_diff,E3CI = E3CI, GDP_diff = GDP, CO2_diff = CO2, GHG_diff = GHG, FA_diff = FA), Year = Year))
}
```

```{r}
select_year_diff2 <- function(year = 2000){
  Year <- NULL
  E3CI_diff <- NULL
  E3CI <- NULL
  CO2 <- NULL
  GDP <- NULL
  GHG <- NULL
  FA <- NULL
  idarea <- NULL
  idtime <- NULL
  # Lo hacemos asi porque no he comprobado que el orden de los códigos sea el   mismo
  # De hecho dudo que lo sea
  for(i in shp_world_filtered_unique$color_code){
    cond <- which((clean_data$Code == i)&(clean_data$Year == year))
    E3CI_diff <- c(E3CI_diff,clean_data[cond,"E3CI_diff"])
    E3CI <- c(E3CI,clean_data[cond,"E3CI"])
    CO2 <- c(CO2, clean_data[cond,"CO2_diff"])
    GDP <- c(GDP, clean_data[cond,"GDP_diff"])
    GHG <- c(GHG, clean_data[cond,"GHG_diff"])
    FA <- c(FA, clean_data[cond,"FA_diff"])
    idarea <- c(idarea, clean_data[cond, "idarea"])
    idtime <- c(idtime, clean_data[cond, "idtime"])
    Year <- c(Year,year)}

  return(cbind(shp_world_filtered_unique,data.frame(E3CI_diff = E3CI_diff,E3CI = E3CI, GDP_diff = GDP, CO2_diff = CO2, GHG_diff = GHG, FA_diff = FA), Year = Year, idarea = idarea, idtime = idtime))
}
```


```{r}
data_2000 <- select_year_diff(2000)
head(data_2000[,c("color_code","E3CI_diff","GDP_diff","CO2_diff","GHG_diff","FA_diff","Year")])
#View(data_2000)
```

#### Map - E3CI
```{r}
map2000 <- ggplot(data = data_2000) +
  geom_sf(aes(fill = E3CI)) + 
  labs(fill = "ERA5 - Risk Index") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red")

map2000_diff <- ggplot(data = data_2000) +
  geom_sf(aes(fill = E3CI_diff)) + 
  labs(fill = "ERA5 - Risk Index") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red")

map2000
map2000_diff
```


#### Interactive Map E3CI

Cooler map

```{r}
library(leaflet)
cool_map2000 <- leaflet(data_2000) %>% addTiles()
pal <- colorNumeric(palette = "YlOrRd", domain = data_2000$E3CI)
cool_map2000 %>% 
  addPolygons(color = "grey", weight = 1, 
              fillColor = ~pal(E3CI), fillOpacity = 0.5) %>%
  addLegend(pal = pal, values = ~E3CI, opacity = 0.5, 
            title = "E3CI", position = "bottomright")
```

```{r}
library(INLA)
```

We are dealing with an irregular lattice. The first step is to create an adjacency matrix (neighbors matrix):

**Neighborhood Matrix**

By default, it will create a binary adjacency matrix, so that two regions are neighbors only if they share at least one point in common boundary (i.e., it is a queen adjacency).
```{r, warning=FALSE}
library(spdep)
nb <- spdep::poly2nb(data_2000, row.names = data_2000$color_code)

plot(data_2000$geometry)
plot(nb, coords = data_2000$geometry, col = "blue", lwd = 2, add = TRUE)
```

```{r}
nb2INLA("map.adj", nb)
g <- inla.read.graph(filename = "map.adj")
```


These are indices (for the areas and years) needed to fit the INLA models:
```{r}
clean_data$idarea <- as.numeric(as.factor(clean_data$Country))
clean_data$idtime <- 1 + clean_data$Year - min(clean_data$Year)
```




