---
title: "Spatio_Temporal_Modelling_3"
author: "Luis Jaime Vidal Jordi"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Warning
**Run the RMD `Cleaning Pre-Processing` before.**

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This RMarkdown contains part of the spatio-temporal modelling code elaborated for the Luis' Master Thesis. To follow the code it is highly advisable to previously read the thesis.

```{r}
# Some Libraries
library(INLA)
library(tidyverse)
library(magrittr)
library(xtable)
library(cowplot)
library(GGally)
```



## SPATIO-TEMPORAL MODELS: Outlier assessment.

So far we have seen that the *BYM* model is able to fit one year's data very well. However, this fit is not valid (at least not as good as one would expect) for predicting different years. Then the idea is clear: try to estimate the time dependence.


The models we are going to consider assume **separability**. That is, the covariance structure can be decomposed as a spatial and a temporal term.

### MODEL NO OUTLIERS
We start by selecting 10 years: from 1990 to 1999.

#### Selecting data
```{r}
years9099 <- 1982:2005
data90_99 <- NULL
for(i in years9099){
  data90_99 <- rbind(data90_99,select_year_diff(year = i))
}
```


#### Raw Correlations

```{r,message=FALSE}
numeric_9099 <- data_frame("E3CI" = data90_99$E3CI_diff,
                           "GDP_diff" = data90_99$GDP_diff,
                           "CO2_diff" = data90_99$CO2_diff,
                           "GHG_diff" = data90_99$GHG_diff,
                           "FA_diff" = data90_99$FA_diff)

cor_9099 <- cor(numeric_9099, use = "complete.obs")

library(ggcorrplot)
ggcorrplot(cor_9099, 
           method = "circle", 
           type = "lower", 
           lab = TRUE, 
           lab_size = 3, 
           colors = c("blue", "white", "red"), 
           title = "Correlogram")

ggpairs(data = data90_99, columns = c("E3CI_diff", "GDP_diff", "CO2_diff","GHG_diff", "FA_diff"))
```

#### Grouped Correlations

```{r}
library(dplyr)
library(ggplot2)
library(sf)
library(patchwork)

# Calculate correlations by group
correlation_by_group <- data90_99 %>%
  group_by(color_code) %>%
  summarise(corrE3CI_GDP = cor(E3CI_diff, GDP_diff),
            corrGDP_CO2 = cor(GDP_diff, CO2_diff),
            corrCO2_GHG = cor(GHG_diff, CO2_diff),
            corrCO2_E3CI = cor(CO2_diff, E3CI_diff))

# Base theme to remove axis labels and grid
base_theme <- theme_minimal() +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank()
  )

# Plot for E3CI and GDP correlation
plot1 <- ggplot(data = correlation_by_group) +
  geom_sf(aes(fill = corrE3CI_GDP)) +
  scale_fill_gradient2(low = "purple", mid = "white", high = "yellow", midpoint = 0,
                       name = "Correlation") +
  labs(title = "E3CI and GDP") +
  base_theme +
  theme(legend.position = "left")

# Plot for CO2 and GDP correlation
plot2 <- ggplot(data = correlation_by_group) +
  geom_sf(aes(fill = corrGDP_CO2)) +
  scale_fill_gradient2(low = "purple", mid = "white", high = "yellow", midpoint = 0,
                       name = "Correlation") +
  labs(title = "CO2 and GDP") +
  base_theme +
  theme(legend.position = "right")

# Plot for CO2 and GHG correlation
plot3 <- ggplot(data = correlation_by_group) +
  geom_sf(aes(fill = corrCO2_GHG)) +
  scale_fill_gradient2(low = "purple", mid = "white", high = "yellow", midpoint = 0,
                       name = "Correlation") +
  labs(title = "CO2 and GHG") +
  base_theme +
  theme(legend.position = "left")

# Plot for CO2 and E3CI correlation
plot4 <- ggplot(data = correlation_by_group) +
  geom_sf(aes(fill = corrCO2_E3CI)) +
  scale_fill_gradient2(low = "purple", mid = "white", high = "yellow", midpoint = 0,
                       name = "Correlation") +
  labs(title = "CO2 and E3CI") +
  base_theme +
  theme(legend.position = "right")

# Combine the four plots into one
combined_plot <- (plot1 + plot2) / (plot3 + plot4)

# Display the combined plot
print(combined_plot)

```

The map that shows the correlation between E3CI and GDP over time could be an indicator of the difficulty to modelize the index with our covariates.

#### 1. Parametric Trend

$$
\eta_{it} = \text{fixed effects } + v_i + \nu_i + (\beta + \delta_i)\times t
$$

```{r}
# We create indexes to use the f() notation
data90_99$idarea <- as.numeric(as.factor(data90_99$color_code))
data90_99$idarea1 <- data90_99$idarea
data90_99$idtime <- 1 + data90_99$Year - min(data90_99$Year)

# (constr = TRUE) is added in some references but result does not change

formula9090_1 <- E3CI_diff ~ 1 + GDP_diff + CO2_diff + FA_diff + 
  f(idarea, model = "bym", graph = g) + 
  f(idarea1,idtime, model = "iid") + idtime
```


```{r}
mod90_99 <- inla(formula = formula9090_1, 
                 family = "gaussian",
                 data = data90_99,
                 control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE),
                 control.predictor = list(compute = TRUE))


```

```{r}
data90_99$fitted <- mod90_99$summary.fitted.values[,"mean"]
```


```{r}
Fitted90_00 <- ggplot(data90_99) + geom_sf(aes(fill = fitted)) +
  facet_wrap(~Year, dir = 'h') +
  ggtitle("Model 1 E3CI") + theme_bw() + 
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()
  ) +
  scale_fill_gradient2(
    midpoint = -0, low = "blue", mid = "white", high = "red"
  )

Real90_00 <- ggplot(data90_99) + geom_sf(aes(fill = E3CI_diff)) +
  facet_wrap(~Year, dir = 'h') +
  ggtitle("Real E3CI") + theme_bw() + 
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()
  ) +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red"
  )
```




#### 2. Non-Parametric Dynamic Trend

$$
\eta_{it} = \text{fixed} + v_i + \nu_i + \gamma_t + \phi_t
$$

here $\gamma_t$ represents a random walk and $\phi_t$ an (prior exchangeable) Gaussian random effect.

```{r}
data90_99$idtime1 <- data90_99$idtime
formula_9090_2 <- E3CI_diff ~ 1 + GDP_diff + CO2_diff + FA_diff + f(idarea, model = "bym", graph = g) + 
  f(idtime, model = "rw1") +
  f(idtime1, model = "iid")

mod90_99_2 <- inla(formula = formula_9090_2, 
                 family = "gaussian",
                 data = data90_99,
                 control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE),
                 control.predictor = list(compute = TRUE))
```

```{r}
data90_99$fitted2 <- mod90_99_2$summary.fitted.values[,"mean"]

Fitted90_00_2 <- ggplot(data90_99) + geom_sf(aes(fill = fitted2)) +
  facet_wrap(~Year, dir = 'h') +
  ggtitle("rw1 E3CI") + theme_bw() + 
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()
  ) +
  scale_fill_gradient2(
    midpoint = -0, low = "blue", mid = "white", high = "red"
  )
```


#### 3-6. Interaction Time-Space

The models we are going to fit in this sectio are an extension of the above one, but now a interaction between time and space is considered through the term: $\delta_{it}$:

$$
\eta_{it} = \text{fixed} + v_i + u_i + \gamma_t + \phi_t + \delta_{it}
$$

Now the point is how to define this interaction. There are several ways to define it. Let us consider 4 types:

- I: Interaction between both unstructured spatial $v_i$ and temporal effects $\phi_t $.

- II: Interaction between unstructured spatial $v_i$ and structured temporal $\gamma_t$.

- III: Interaction between structured spatial $u_i$ and unstructured temporal $\phi_t$.

- IV: Interaction between both structured spatial $u_i$ and temporal $\gamma_t$ effects.


**Model I**

Creating the interaction term:
```{r}
data90_99$idareatime <- interaction(data90_99$idarea,data90_99$idtime)
data90_99$idareatime <- as.numeric(data90_99$idareatime)
```

```{r}
formula_9090_I <- E3CI_diff ~ 1 + GDP_diff + CO2_diff + FA_diff + 
  f(idarea, model = "bym", graph = g) + 
  f(idtime, model = "rw1") + f(idtime1, model = "iid") +
  f(idareatime, model = "iid")
```

```{r}
mod90_99_I <- inla(formula = formula_9090_I, 
                 family = "gaussian",
                 data = data90_99,
                 control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE),
                 control.predictor = list(compute = TRUE, link = 1))

data90_99$fittedI <- mod90_99_I$summary.fitted.values[,"mean"]

Fitted90_00_I <- ggplot(data90_99) + geom_sf(aes(fill = fittedI)) +
  facet_wrap(~Year, dir = 'h') +
  ggtitle("Interaction E3CI") + theme_bw() + 
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()
  ) +
  scale_fill_gradient2(
    midpoint = -0, low = "blue", mid = "white", high = "red"
  )
```


**Model II**

To specify this model we need to create one more copy for the time and area indexes:

```{r}
data90_99$idarea_int <- data90_99$idarea
data90_99$idtime_int <- data90_99$idtime
```


```{r}
formula_9090_II <- E3CI_diff ~ 1 + GDP_diff + CO2_diff + FA_diff + 
  f(idarea, model = "bym", graph = g) + 
  f(idtime, model = "rw1") + f(idtime1, model = "iid") +
  f(idarea_int, model = "iid",group = idtime_int,
    control.group = list(model="rw1"))
```

```{r}
mod90_99_II <- inla(formula = formula_9090_II, 
                 family = "gaussian",
                 data = data90_99,
                 control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE),
                 control.predictor = list(compute = TRUE))

data90_99$fittedII <- mod90_99_II$summary.fitted.values[,"mean"]

Fitted90_00_II <- ggplot(data90_99) + geom_sf(aes(fill = fittedII)) +
  facet_wrap(~Year, dir = 'h') +
  ggtitle("Interaction E3CI") + theme_bw() + 
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()
  ) +
  scale_fill_gradient2(
    midpoint = -0, low = "blue", mid = "white", high = "red"
  )
```

**Model III**

```{r}
formula_9090_III <- E3CI_diff ~ 1 + GDP_diff + CO2_diff + FA_diff + 
  f(idarea, model = "bym", graph = g) + 
  f(idtime, model = "rw1") + 
  f(idtime1, model = "iid") +
  f(idtime_int, model = "iid",group = idarea_int,
    control.group = list(model="besag", graph = g))
```

```{r}
mod90_99_III <- inla(formula = formula_9090_III, 
                 family = "gaussian",
                 data = data90_99,
                 control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE),
                 control.predictor = list(compute = TRUE))

data90_99$fittedIII <- mod90_99_III$summary.fitted.values[,"mean"]

Fitted90_00_III <- ggplot(data90_99) + geom_sf(aes(fill = fittedIII)) +
  facet_wrap(~Year, dir = 'h') +
  ggtitle("Interaction E3CI") + theme_bw() + 
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()
  ) +
  scale_fill_gradient2(
    midpoint = -0, low = "blue", mid = "white", high = "red"
  )
```

**Model IV**

The most complex interaction, as we are going to assume interaction between the structured effects:
```{r}
formula_9090_IV <- E3CI_diff ~ 1 + GDP_diff + CO2_diff + FA_diff + 
  f(idarea, model = "bym", graph = g) + 
  f(idtime, model = "rw1") + 
  f(idtime1, model = "iid") +
  f(idarea_int, model = "besag", graph = g,
    group = idtime_int,
    control.group = list(model="rw1"))
```

```{r}
mod90_99_IV <- inla(formula = formula_9090_IV, 
                 family = "gaussian",
                 data = data90_99,
                 control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE),
                 control.predictor = list(compute = TRUE))

data90_99$fittedIV <- mod90_99_IV$summary.fitted.values[,"mean"]

Fitted90_00_IV <- ggplot(data90_99) + geom_sf(aes(fill = fittedIV)) +
  facet_wrap(~Year, dir = 'h') +
  ggtitle("Interaction E3CI") + theme_bw() + 
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()
  ) +
  scale_fill_gradient2(
    midpoint = -0, low = "blue", mid = "white", high = "red"
  )
```


#### Model Comparison
For the first three spatio-temporal models we cannot expect good results, as we have seen with the spatial models and correlations that the time-space-covariance relationship is not so simple.

- Basic Spatio-Temporal Models:
```{r}
# Real90_00
# Fitted90_00
# Fitted90_00_2
```


- More complex spatio-temporal models: interactions
```{r}
# Real90_00
# Fitted90_00_I
# Fitted90_00_II
# Fitted90_00_III
# Fitted90_00_IV
```

Combine these plots

```{r}
# comb_int1_real <- plot_grid(Real90_00,Fitted90_00_I,ncol = 2) %>% print()
# comb_paramtrend_real <- plot_grid(Real90_00,Fitted90_00,ncol = 2) %>% print()
# comb_int4_real <- plot_grid(Real90_00,Fitted90_00_IV,ncol = 2) %>% print()
```



```{r}
results9099 <- data.frame("E3CI_real" = data90_99$E3CI_diff,"IID" = mod90_99$summary.fitted.values[,"mean"], "rw1" = mod90_99_2$summary.fitted.values[,"mean"], "Interaction_I" = mod90_99_I$summary.fitted.values[,"mean"],"Interaction_II" = mod90_99_II$summary.fitted.values[,"mean"],"Interaction_III" = mod90_99_III$summary.fitted.values[,"mean"],"Interaction_IV" = mod90_99_IV$summary.fitted.values[,"mean"])

mse <- function(real,fitted){
  return(mean((real - fitted)^2))
}
results_df9099 <- cbind(data90_99, results9099)
```

Different available criteria for model selection:

- `MSE` (on train partition)
```{r}
MSE_90_99 <- c(mse(results_df9099$E3CI_diff, fitted = results9099$E3CI_real),
               mse(results_df9099$E3CI_diff, fitted = results9099$IID), 
               mse(results_df9099$E3CI_diff, fitted = results9099$rw1), 
               mse(results_df9099$E3CI_diff, fitted = results9099$Interaction_I),
               mse(results_df9099$E3CI_diff, fitted = results9099$Interaction_II),
               mse(results_df9099$E3CI_diff, fitted = results9099$Interaction_III),
               mse(results_df9099$E3CI_diff, fitted = results9099$Interaction_IV))
#MSE_90_99
```


- `cpo` can be computed in the following way:
```{r}
cpo <- function(inla_model){
  return(-sum(log(inla_model$cpo$cpo)))
}

cpo_iid_time <- cpo(mod90_99)
cpo_rw1 <- cpo(mod90_99_2)
cpo_interactionI <- cpo(mod90_99_I)
cpo_interactionII <- cpo(mod90_99_II)
cpo_interactionIII <- cpo(mod90_99_III)
cpo_interactionIV <- cpo(mod90_99_IV)


CPOs_9099 <- c(cpo_iid_time,cpo_rw1,cpo_interactionI,cpo_interactionII,
               cpo_interactionIII,cpo_interactionIV)
```

- `marginal likelihood`: **Hay dos valores, no se cual es el que pone la teoria**
```{r}
MLIKs9099 <- c(mod90_99$mlik[[1,1]],mod90_99_2$mlik[[1,1]],
               mod90_99_I$mlik[[1,1]],
               mod90_99_II$mlik[[1,1]],
               mod90_99_III$mlik[[1,1]],
               mod90_99_IV$mlik[[1,1]])
```

- `DIC`
```{r}
DICs9099 <- c(mod90_99$dic$dic,mod90_99_2$dic$dic,
              mod90_99_I$dic$dic,
              mod90_99_II$dic$dic,
              mod90_99_III$dic$dic,
              mod90_99_IV$dic$dic)
```

- `WAIC` 
```{r}
WAICs9099 <- c(mod90_99$waic$waic,mod90_99_2$waic$waic,
               mod90_99_I$waic$waic,
               mod90_99_II$waic$waic,
               mod90_99_III$waic$waic,
               mod90_99_IV$waic$waic)
```


```{r}
percentage_diff_sign <- function(vector1, vector2) {
  # Check that the vectors have the same length
  if (length(vector1) != length(vector2)) {
    stop("The vectors must have the same length.")
  }
  
  # Calculate the number of elements with differing signs
  diff_signs <- sum(sign(vector1) != sign(vector2))
  
  # Calculate the percentage
  percentage <- (diff_signs / length(vector1)) * 100
  
  return(percentage)
}
signs90_99 <- c(percentage_diff_sign(results_df9099$E3CI_diff,results_df9099$fitted),
                percentage_diff_sign(results_df9099$E3CI_diff,results_df9099$fitted2),
                percentage_diff_sign(results_df9099$E3CI_diff,results_df9099$fittedI),
                percentage_diff_sign(results_df9099$E3CI_diff,results_df9099$fittedII),
                percentage_diff_sign(results_df9099$E3CI_diff,results_df9099$fittedIII),
                percentage_diff_sign(results_df9099$E3CI_diff,results_df9099$fittedIV))
```

```{r}
errors_df9099_n <- data.frame("Likelihood" = rep("gaussian",6), 
                              "Model" = c("bym + IID", "bym + rw1", "Interaction I",
                                        "Interaction II","Interaction III","Interaction IV"), 
               "WAIC" = WAICs9099,
               "MLIK" = MLIKs9099,
               "CPO" = CPOs_9099,
               "DIC" = DICs9099,
               "MSE" = MSE_90_99[2:length(MSE_90_99)],
               "SignError" = signs90_99)

library(kableExtra)
kable(errors_df9099_n)
#xtable::xtable(errors_df9099_n, digits = 3)
```

#### Predictions - No outliers 

First, we save the current `data90_99` which contains the fitted values of the above models
```{r}
data90_99_saved <- data90_99
```

To predict with INLA we only have to include the observations we want to predict **with a missing value in the response**. As we are setting from 2014 to 2022 as NA there are no outliers in the train set. 

Then, let us reconsider `data_9099` but now we are going to include the year 2022:
```{r}
years9099 <- c(1982:2022)
data90_99 <- NULL
for(i in years9099){ 
  data90_99 <- rbind(data90_99,select_year_diff2(year = i))
}
data90_99_bench <- data90_99
data90_99[which(data90_99$Year %in% c(2021:2022)),"E3CI_diff"] = NA
```

Now, we refit the best model obtained above: *bym + interaction I*.

```{r}
# Create indexes as before
#data90_99$idarea <- as.numeric(as.factor(data90_99$color_code))
data90_99$idarea1 <- data90_99$idarea
#data90_99$idtime <- 1 + data90_99$Year - min(data90_99$Year)
data90_99$idtime1 <- data90_99$idtime
data90_99$idareatime <- interaction(data90_99$idarea,data90_99$idtime)
data90_99$idtime_int <- data90_99$idtime
data90_99$idarea_int <- data90_99$idarea

formula_9090_I <- E3CI_diff ~ 1 + GDP_diff + CO2_diff + FA_diff + 
  f(idarea, model = "bym", graph = g) + 
  f(idtime, model = "rw2") + f(idtime1, model = "iid") +
  f(idareatime, model = "iid")

# Fit the model
mod90_99_predI <- inla(formula = formula_9090_I, 
                 family = "gaussian",
                 data = data90_99,
                 control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE),
                 control.predictor = list(compute = TRUE, link = 1))
```

Now, Inla has used the predictive distribution to compute the values for the year 2000:

```{r}
rows2000 <- which(data90_99$Year == 2000)
rows2000_bench <- which(data90_99_bench$Year == 2000)
pred_2000 <- mod90_99_predI$summary.fitted.values[rows2000,"mean"]

#preds_mod9099 <- data.frame("Real_2000" = data90_99_bench[rows2000_bench,"E3CI"])
data90_99_bench$fittedI <- mod90_99_predI$summary.fitted.values[,"mean"]
data90_99_bench <- data90_99_bench %>%  relocate(fittedI, .after = E3CI)
```

Then, we can check the MSE error of the predicted values:

```{r}
# To check whether the order is correct or not 
unique(rows2000_bench == rows2000)
unique(data90_99$color_code[rows2000] == data90_99_bench$color_code[rows2000_bench])


df_pred90_99 <- data.frame("Real" = data90_99_bench$E3CI_diff[rows2000_bench],
                      "fittedI" = data90_99_bench$fittedI[rows2000_bench],
                      "Code1" = data90_99_bench$color_code[rows2000_bench],
                      "Code2" = data90_99$color_code[rows2000_bench])

#View(df_pred90_99)

MSE_pred2000 <- mse(real = data90_99_bench$E3CI_diff[rows2000_bench],
                    fitted = data90_99_bench$fittedI[rows2000_bench]) %>% print()
```

It could be also of interest to check the number of missclassified signs:
```{r}
percentage_diff_sign <- function(vector1, vector2) {
  # Check that the vectors have the same length
  if (length(vector1) != length(vector2)) {
    stop("The vectors must have the same length.")
  }
  
  # Calculate the number of elements with differing signs
  diff_signs <- sum(sign(vector1) != sign(vector2))
  
  # Calculate the percentage
  percentage <- (diff_signs / length(vector1)) * 100
  
  return(percentage)
}
```


```{r}
map2000_real <- ggplot(data = data90_99_bench[rows2000_bench,]) +
  geom_sf(aes(fill = E3CI_diff)) + 
  labs(fill = "Real ERA5 - 2000") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red")

map2000_predI <- ggplot(data = data90_99_bench[rows2000_bench,]) +
  geom_sf(aes(fill = fittedI)) + 
  labs(fill = "Bym+Interaction ERA5 - 2000") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red")

map2000_real
map2000_predI
Mse_pred2000 <- mse(data90_99_bench$E3CI_diff[rows2000_bench],data90_99_bench$fittedI[rows2000_bench])

signerror2000<- percentage_diff_sign(data90_99_bench$E3CI_diff[rows2000_bench],data90_99_bench$fittedI[rows2000_bench])
```

```{r}
rows2001_bench <- which(data90_99_bench$Year == 2001)
map2001_real <- ggplot(data = data90_99_bench[rows2001_bench,]) +
  geom_sf(aes(fill = E3CI_diff)) + 
  labs(fill = "Real ERA5 - 2001") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red")

map2001_predI <- ggplot(data = data90_99_bench[rows2001_bench,]) +
  geom_sf(aes(fill = fittedI)) + 
  labs(fill = "Bym+Interaction ERA5 - 2001") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red")

map2001_real
map2001_predI
Mse_pred2001 <- mse(data90_99_bench$E3CI_diff[rows2001_bench],data90_99_bench$fittedI[rows2001_bench]) %>% print()

signerror2001 <- percentage_diff_sign(data90_99_bench$E3CI_diff[rows2001_bench],data90_99_bench$fittedI[rows2001_bench]) %>% print()
```


```{r}
rows2002_bench <- which(data90_99_bench$Year == 2002)
map2002_real <- ggplot(data = data90_99_bench[rows2002_bench,]) +
  geom_sf(aes(fill = E3CI_diff)) + 
  labs(fill = "Real ERA5 - 2002") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red")

map2002_predI <- ggplot(data = data90_99_bench[rows2002_bench,]) +
  geom_sf(aes(fill = fittedI)) + 
  labs(fill = "Bym+Interaction ERA5 - 2002") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red")

map2002_real
map2002_predI

Mse_pred2002 <- mse(data90_99_bench$E3CI_diff[rows2002_bench],data90_99_bench$fittedI[rows2002_bench])

signerror2002 <- percentage_diff_sign(data90_99_bench$E3CI_diff[rows2002_bench],data90_99_bench$fittedI[rows2002_bench]) %>% print()
```

```{r}
rows2003_bench <- which(data90_99_bench$Year == 2003)
map2003_real <- ggplot(data = data90_99_bench[rows2003_bench,]) +
  geom_sf(aes(fill = E3CI_diff)) + 
  labs(fill = "Real ERA5 - 2003") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red")

map2003_predI <- ggplot(data = data90_99_bench[rows2003_bench,]) +
  geom_sf(aes(fill = fittedI)) + 
  labs(fill = "Bym+Interaction ERA5 - 2003") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red")

map2003_real
map2003_predI
Mse_pred2003 <- mse(data90_99_bench$E3CI_diff[rows2003_bench],data90_99_bench$fittedI[rows2003_bench])
signerror2003 <- percentage_diff_sign(data90_99_bench$E3CI_diff[rows2003_bench],data90_99_bench$fittedI[rows2003_bench]) %>% print()
```

```{r}
rows2004_bench <- which(data90_99_bench$Year == 2004)
map2004_real <- ggplot(data = data90_99_bench[rows2004_bench,]) +
  geom_sf(aes(fill = E3CI_diff)) + 
  labs(fill = "Real ERA5 - 2004") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red")

map2004_predI <- ggplot(data = data90_99_bench[rows2004_bench,]) +
  geom_sf(aes(fill = fittedI)) + 
  labs(fill = "Bym+Interaction ERA5 - 2004") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red")

map2004_real
map2004_predI

Mse_pred2004 <- mse(data90_99_bench$E3CI_diff[rows2004_bench],data90_99_bench$fittedI[rows2004_bench])

signerror2004 <- percentage_diff_sign(data90_99_bench$E3CI_diff[rows2004_bench],data90_99_bench$fittedI[rows2004_bench]) %>% print()
```


```{r}
rows2005_bench <- which(data90_99_bench$Year == 2005)
map2005_real <- ggplot(data = data90_99_bench[rows2005_bench,]) +
  geom_sf(aes(fill = E3CI_diff)) + 
  labs(fill = "Real ERA5 - 2005") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red")

map2005_predI <- ggplot(data = data90_99_bench[rows2005_bench,]) +
  geom_sf(aes(fill = fittedI)) + 
  labs(fill = "Bym+Interaction ERA5 - 2005") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red")

map2005_real
map2005_predI
Mse_pred2005 <- mse(data90_99_bench$E3CI_diff[rows2005_bench],data90_99_bench$fittedI[rows2005_bench])

signerror2005 <- percentage_diff_sign(data90_99_bench$E3CI_diff[rows2005_bench],data90_99_bench$fittedI[rows2005_bench]) %>% print()
```

All joint:

```{r}
filtered_data <- data90_99_bench %>%
  filter(Year %in% c(2001, 2002, 2003, 2004, 2005))

combined_df <- filtered_data %>%
  dplyr::select(geometry, Year, E3CI_diff, fittedI) %>%
  pivot_longer(cols = c(E3CI_diff, fittedI), 
               names_to = "Variable", 
               values_to = "Value") %>%
  unite("Year_Variable", Year, Variable, sep = "_")

combined_map <- ggplot(combined_df) +
  geom_sf(aes(fill = Value)) + 
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red") +
  facet_wrap(~ Year_Variable, ncol = 5) +
  labs(fill = "Value") +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        strip.background = element_blank(),
        strip.text = element_text(face = "bold", size = 12),
        legend.position = "right") 

print(combined_map)
```

In a table
```{r}
preds_df9099 <- data.frame(Year = 2000:2005,
                           SignError = c(signerror2000,signerror2001,
                                          signerror2002,signerror2003,
                                          signerror2004,signerror2005),
                           MSE = c(Mse_pred2000,Mse_pred2001,
                                    Mse_pred2002,Mse_pred2003,
                                    Mse_pred2004,Mse_pred2005))

# transformed_df <- data.frame(
#   Metric = c("SignError","SignError","MSE","MSE"),
#   Model = c("Spatial", "Spatio-Temp","Spatial", "Spatio-Temp"),
#   `2000` = c(signerror2000_00, signerror2000, MSE_2000_00, Mse_pred2000),
#   `2001` = c(signerror2001_00, signerror2001, MSE_2001_00, Mse_pred2001),
#   `2002` = c(signerror2002_00, signerror2002, MSE_2002_00, Mse_pred2002),
#   `2003` = c(signerror2003_00, signerror2003, MSE_2003_00, Mse_pred2003),
#   `2004` = c(signerror2004_00, signerror2004, MSE_2004_00, Mse_pred2004),
#   `2005` = c(signerror2005_00, signerror2005, MSE_2005_00, Mse_pred2005)
# )

transformed_df <- data.frame(
  Metric = c("SignError","SignError","MSE","MSE"),
  Model = c("Spatial", "Spatio-Temp","Spatial", "Spatio-Temp"),
  `2000` = c(NA, signerror2000, NA, Mse_pred2000),
  `2001` = c(NA, signerror2001, NA, Mse_pred2001),
  `2002` = c(NA, signerror2002, NA, Mse_pred2002),
  `2003` = c(NA, signerror2003, NA, Mse_pred2003),
  `2004` = c(NA, signerror2004, NA, Mse_pred2004),
  `2005` = c(NA, signerror2005, NA, Mse_pred2005)
)

kable(transformed_df)
xtable(transformed_df, digits = 3)
```

##### All the sign errors
```{r}
compute_summary <- function(years){
  sign_errors <- NULL
  mses <- NULL
  for(i in years){
    cond <- which(data90_99_bench$Year == i)
    sign_errors <- c(sign_errors,
                     percentage_diff_sign(data90_99_bench$E3CI_diff[cond],
                                          data90_99_bench$fittedI[cond]))
    mses <- c(mses, 
              mse(data90_99_bench$E3CI_diff[cond],
                  data90_99_bench$fittedI[cond]))
  }
  df <- data.frame(Year = years,
                   SignError = sign_errors,
                   MSE = mses)
  return(df)
}
```

```{r}
del82_05_202122 <- compute_summary(years9099)
kable(compute_summary(years9099))

```


De momento los años mas problematicos son 2014, 2018 y 2020.

```{r}
outliers <- which((data90_99_bench$E3CI_diff >= 4) | (data90_99_bench$E3CI_diff <= -4))
outliers_GDP <- which((data90_99_bench$GDP_diff >= 4) | (data90_99_bench$GDP_diff <= -4))
#View(data90_99_bench[outliers,c("name","E3CI_diff","fittedI","Year","GDP_diff","CO2_diff","FA_diff")])
```

##### 2014, 2020, 2021 and 2022

```{r}
rows2014_bench_e <- which(data90_99_bench$Year == 2014)
map2014_real_e <- ggplot(data = data90_99_bench[rows2014_bench_e,]) +
  geom_sf(aes(fill = E3CI_diff)) + 
  labs(fill = "Real - 2014") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red") + 
  theme(legend.position = "left",
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA)
  )

map2014_predI_e <- ggplot(data = data90_99_bench[rows2014_bench_e,]) +
  geom_sf(aes(fill = fittedI)) + 
  labs(fill = "Interaction I - 2014") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red") + 
  theme(legend.position = "left",
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA)
  )

map2014_real_e
map2014_predI_e
Mse_pred2014 <- mse(data90_99_bench$E3CI_diff[rows2014_bench_e],data90_99_bench$fittedI[rows2014_bench_e]) %>% print()

signerror2014 <- percentage_diff_sign(data90_99_bench$E3CI_diff[rows2014_bench_e],data90_99_bench$fittedI[rows2014_bench_e]) %>% print()
```



```{r}
rows2020_bench_e <- which(data90_99_bench$Year == 2020)
map2020_real_e <- ggplot(data = data90_99_bench[rows2020_bench_e,]) +
  geom_sf(aes(fill = E3CI_diff)) + 
  labs(fill = "Real - 2020") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red") + 
  theme(legend.position = "top",
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA)
  )

map2020_predI_e <- ggplot(data = data90_99_bench[rows2020_bench_e,]) +
  geom_sf(aes(fill = fittedI)) + 
  labs(fill = "Interaction I - 2020") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red") + 
  theme(legend.position = "bottom",
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA)
  )

map2020_real_e
map2020_predI_e
Mse_pred2020 <- mse(data90_99_bench$E3CI_diff[rows2020_bench_e],data90_99_bench$fittedI[rows2020_bench_e]) %>% print()

signerror2020 <- percentage_diff_sign(data90_99_bench$E3CI_diff[rows2020_bench_e],data90_99_bench$fittedI[rows2020_bench_e]) %>% print()
```

```{r}
rows2021_bench_e <- which(data90_99_bench$Year == 2021)
map2021_real_e <- ggplot(data = data90_99_bench[rows2021_bench_e,]) +
  geom_sf(aes(fill = E3CI_diff)) + 
  labs(fill = "Real - 2021") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red") + 
  theme(legend.position = "right",
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA)
  )

map2021_predI_e <- ggplot(data = data90_99_bench[rows2021_bench_e,]) +
  geom_sf(aes(fill = fittedI)) + 
  labs(fill = "Interaction I - 2021") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red") + 
  theme(legend.position = "right",
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA)
  )

map2021_real_e
map2021_predI_e
Mse_pred2021 <- mse(data90_99_bench$E3CI_diff[rows2021_bench_e],data90_99_bench$fittedI[rows2021_bench_e]) %>% print()

signerror2021 <- percentage_diff_sign(data90_99_bench$E3CI_diff[rows2021_bench_e],data90_99_bench$fittedI[rows2021_bench_e]) %>% print()
```

```{r}
rows2022_bench_e <- which(data90_99_bench$Year == 2022)

map2022_real_e <- ggplot(data = data90_99_bench[rows2022_bench_e,]) +
  geom_sf(aes(fill = E3CI_diff)) + 
  labs(fill = "Real - 2022") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red") + 
  theme(legend.position = "right",
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA)
  )

map2022_predI_e <- ggplot(data = data90_99_bench[rows2022_bench_e,]) +
  geom_sf(aes(fill = fittedI)) + 
  labs(fill = "Interaction I - 2022") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red") + 
  theme(legend.position = "right",
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA)
  )

map2022_real_e
map2022_predI_e
Mse_pred2022 <- mse(data90_99_bench$E3CI_diff[rows2022_bench_e],data90_99_bench$fittedI[rows2022_bench_e]) %>% print()

signerror2022 <- percentage_diff_sign(data90_99_bench$E3CI_diff[rows2022_bench_e],data90_99_bench$fittedI[rows2022_bench_e]) %>% print()
```



```{r}
comb_plot_202122 <- (map2014_real_e +
                       map2020_real_e +
                       map2021_real_e)/(map2014_predI_e +
                                          map2020_predI_e +
                                          map2021_predI_e)



print(comb_plot_202122)
```

#### Posterior Information

```{r}
#summary(mod90_99_predI)
mod90_99_predI$summary.fixed[,1:6]
mod90_99_predI$summary.hyperpar

xtable(mod90_99_predI$summary.fixed[,1:6], digits = 4)
xtable(mod90_99_predI$summary.hyperpar)
```


### MODEL OUTLIERS - Predict 2021?
The same model than above is fitted but setting a t-Student likelihood.

#### Selecting data
```{r}
years9099 <- 1982:2022
data90_99 <- NULL
for(i in years9099){
  data90_99 <- rbind(data90_99,select_year_diff2(year = i))
}
```


#### Raw Correlations

```{r,message=FALSE}
numeric_9099 <- data_frame("E3CI" = data90_99$E3CI_diff,
                           "GDP_diff" = data90_99$GDP_diff,
                           "CO2_diff" = data90_99$CO2_diff,
                           "GHG_diff" = data90_99$GHG_diff,
                           "FA_diff" = data90_99$FA_diff)

cor_9099 <- cor(numeric_9099, use = "complete.obs")

library(ggcorrplot)
ggcorrplot(cor_9099, 
           method = "circle", 
           type = "lower", 
           lab = TRUE, 
           lab_size = 3, 
           colors = c("blue", "white", "red"), 
           title = "Correlogram")

ggpairs(data = data90_99, columns = c("E3CI_diff", "GDP_diff", "CO2_diff","GHG_diff", "FA_diff"))
```

#### Grouped Correlations

```{r}
library(dplyr)
library(ggplot2)
library(sf)
library(patchwork)

# Calculate correlations by group
correlation_by_group <- data90_99 %>%
  group_by(color_code) %>%
  summarise(corrE3CI_GDP = cor(E3CI_diff, GDP_diff),
            corrGDP_CO2 = cor(GDP_diff, CO2_diff),
            corrCO2_GHG = cor(GHG_diff, CO2_diff),
            corrCO2_E3CI = cor(CO2_diff, E3CI_diff))

# Base theme to remove axis labels and grid
base_theme <- theme_minimal() +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank()
  )

# Plot for E3CI and GDP correlation
plot1 <- ggplot(data = correlation_by_group) +
  geom_sf(aes(fill = corrE3CI_GDP)) +
  scale_fill_gradient2(low = "purple", mid = "white", high = "yellow", midpoint = 0,
                       name = "Correlation") +
  labs(title = "E3CI and GDP") +
  base_theme +
  theme(legend.position = "left")

# Plot for CO2 and GDP correlation
plot2 <- ggplot(data = correlation_by_group) +
  geom_sf(aes(fill = corrGDP_CO2)) +
  scale_fill_gradient2(low = "purple", mid = "white", high = "yellow", midpoint = 0,
                       name = "Correlation") +
  labs(title = "CO2 and GDP") +
  base_theme +
  theme(legend.position = "right")

# Plot for CO2 and GHG correlation
plot3 <- ggplot(data = correlation_by_group) +
  geom_sf(aes(fill = corrCO2_GHG)) +
  scale_fill_gradient2(low = "purple", mid = "white", high = "yellow", midpoint = 0,
                       name = "Correlation") +
  labs(title = "CO2 and GHG") +
  base_theme +
  theme(legend.position = "left")

# Plot for CO2 and E3CI correlation
plot4 <- ggplot(data = correlation_by_group) +
  geom_sf(aes(fill = corrCO2_E3CI)) +
  scale_fill_gradient2(low = "purple", mid = "white", high = "yellow", midpoint = 0,
                       name = "Correlation") +
  labs(title = "CO2 and E3CI") +
  base_theme +
  theme(legend.position = "right")

# Combine the four plots into one
combined_plot <- (plot1 + plot2) / (plot3 + plot4)

# Display the combined plot
print(combined_plot)

```

The map that shows the correlation between E3CI and GDP over time could be an indicator of the difficulty to modelize the index with our covariates.

#### 1. Parametric Trend

$$
\eta_{it} = \text{fixed effects } + v_i + \nu_i + (\beta + \delta_i)\times t
$$

```{r}
# We create indexes to use the f() notation
#data90_99$idarea <- as.numeric(as.factor(data90_99$color_code))
data90_99$idarea1 <- data90_99$idarea
#data90_99$idtime <- 1 + data90_99$Year - min(data90_99$Year)

# (constr = TRUE) is added in some references but result does not change

formula9090_1 <- E3CI_diff ~ 1 + GDP_diff + CO2_diff + FA_diff + 
  f(idarea, model = "bym", graph = g) + 
  f(idarea1,idtime, model = "iid") + idtime
```


```{r}
mod90_99 <- inla(formula = formula9090_1, 
                 family = "gaussian",
                 data = data90_99,
                 control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE),
                 control.predictor = list(compute = TRUE))


```

```{r}
data90_99$fitted <- mod90_99$summary.fitted.values[,"mean"]
```


```{r}
Fitted90_00 <- ggplot(data90_99) + geom_sf(aes(fill = fitted)) +
  facet_wrap(~Year, dir = 'h') +
  ggtitle("Model 1 E3CI") + theme_bw() + 
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()
  ) +
  scale_fill_gradient2(
    midpoint = -0, low = "blue", mid = "white", high = "red"
  )

Real90_00 <- ggplot(data90_99) + geom_sf(aes(fill = E3CI_diff)) +
  facet_wrap(~Year, dir = 'h') +
  ggtitle("Real E3CI") + theme_bw() + 
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()
  ) +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red"
  )
```




#### 2. Non-Parametric Dynamic Trend

$$
\eta_{it} = \text{fixed} + v_i + \nu_i + \gamma_t + \phi_t
$$

here $\gamma_t$ represents a random walk and $\phi_t$ an (prior exchangeable) Gaussian random effect.

```{r}
data90_99$idtime1 <- data90_99$idtime
formula_9090_2 <- E3CI_diff ~ 1 + GDP_diff + CO2_diff + FA_diff + f(idarea, model = "bym", graph = g) + 
  f(idtime, model = "rw1") +
  f(idtime1, model = "iid")

mod90_99_2 <- inla(formula = formula_9090_2, 
                 family = "gaussian",
                 data = data90_99,
                 control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE),
                 control.predictor = list(compute = TRUE))
```

```{r}
data90_99$fitted2 <- mod90_99_2$summary.fitted.values[,"mean"]

Fitted90_00_2 <- ggplot(data90_99) + geom_sf(aes(fill = fitted2)) +
  facet_wrap(~Year, dir = 'h') +
  ggtitle("rw1 E3CI") + theme_bw() + 
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()
  ) +
  scale_fill_gradient2(
    midpoint = -0, low = "blue", mid = "white", high = "red"
  )
```


#### 3-6. Interaction Time-Space

The models we are going to fit in this sectio are an extension of the above one, but now a interaction between time and space is considered through the term: $\delta_{it}$:

$$
\eta_{it} = \text{fixed} + v_i + u_i + \gamma_t + \phi_t + \delta_{it}
$$

Now the point is how to define this interaction. There are several ways to define it. Let us consider 4 types:

- I: Interaction between both unstructured spatial $v_i$ and temporal effects $\phi_t $.

- II: Interaction between unstructured spatial $v_i$ and structured temporal $\gamma_t$.

- III: Interaction between structured spatial $u_i$ and unstructured temporal $\phi_t$.

- IV: Interaction between both structured spatial $u_i$ and temporal $\gamma_t$ effects.


**Model I**

Creating the interaction term:
```{r}
data90_99$idareatime <- interaction(data90_99$idarea,data90_99$idtime)
data90_99$idareatime <- as.numeric(data90_99$idareatime)
```

```{r}
formula_9090_I <- E3CI_diff ~ 1 + GDP_diff + CO2_diff + FA_diff + 
  f(idarea, model = "bym", graph = g) + 
  f(idtime, model = "rw1") + f(idtime1, model = "iid") +
  f(idareatime, model = "iid")
```

```{r}
mod90_99_I <- inla(formula = formula_9090_I, 
                 family = "gaussian",
                 data = data90_99,
                 control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE),
                 control.predictor = list(compute = TRUE))

data90_99$fittedI <- mod90_99_I$summary.fitted.values[,"mean"]

Fitted90_00_I <- ggplot(data90_99) + geom_sf(aes(fill = fittedI)) +
  facet_wrap(~Year, dir = 'h') +
  ggtitle("Interaction E3CI") + theme_bw() + 
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()
  ) +
  scale_fill_gradient2(
    midpoint = -0, low = "blue", mid = "white", high = "red"
  )
```


**Model II**

To specify this model we need to create one more copy for the time and area indexes:

```{r}
data90_99$idarea_int <- data90_99$idarea
data90_99$idtime_int <- data90_99$idtime
```


```{r}
formula_9090_II <- E3CI_diff ~ 1 + GDP_diff + CO2_diff + FA_diff + 
  f(idarea, model = "bym", graph = g) + 
  f(idtime, model = "rw1") + f(idtime1, model = "iid") +
  f(idarea_int, model = "iid",group = idtime_int,
    control.group = list(model="rw1"))
```

```{r}
mod90_99_II <- inla(formula = formula_9090_II, 
                 family = "gaussian",
                 data = data90_99,
                 control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE),
                 control.predictor = list(compute = TRUE))

data90_99$fittedII <- mod90_99_II$summary.fitted.values[,"mean"]

Fitted90_00_II <- ggplot(data90_99) + geom_sf(aes(fill = fittedII)) +
  facet_wrap(~Year, dir = 'h') +
  ggtitle("Interaction E3CI") + theme_bw() + 
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()
  ) +
  scale_fill_gradient2(
    midpoint = -0, low = "blue", mid = "white", high = "red"
  )
```

**Model III**

```{r}
formula_9090_III <- E3CI_diff ~ 1 + GDP_diff + CO2_diff + FA_diff + 
  f(idarea, model = "bym", graph = g) + 
  f(idtime, model = "rw1") + 
  f(idtime1, model = "iid") +
  f(idtime_int, model = "iid",group = idarea_int,
    control.group = list(model="besag", graph = g))
```

```{r}
mod90_99_III <- inla(formula = formula_9090_III, 
                 family = "gaussian",
                 data = data90_99,
                 control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE),
                 control.predictor = list(compute = TRUE))

data90_99$fittedIII <- mod90_99_III$summary.fitted.values[,"mean"]

Fitted90_00_III <- ggplot(data90_99) + geom_sf(aes(fill = fittedIII)) +
  facet_wrap(~Year, dir = 'h') +
  ggtitle("Interaction E3CI") + theme_bw() + 
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()
  ) +
  scale_fill_gradient2(
    midpoint = -0, low = "blue", mid = "white", high = "red"
  )
```

**Model IV**

The most complex interaction, as we are going to assume interaction between the structured effects:
```{r}
formula_9090_IV <- E3CI_diff ~ 1 + GDP_diff + CO2_diff + FA_diff + 
  f(idarea, model = "bym", graph = g) + 
  f(idtime, model = "rw1") + 
  f(idtime1, model = "iid") +
  f(idarea_int, model = "besag", graph = g,
    group = idtime_int,
    control.group = list(model="rw1"))
```

```{r}
mod90_99_IV <- inla(formula = formula_9090_IV, 
                 family = "gaussian",
                 data = data90_99,
                 control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE),
                 control.predictor = list(compute = TRUE))

data90_99$fittedIV <- mod90_99_IV$summary.fitted.values[,"mean"]

Fitted90_00_IV <- ggplot(data90_99) + geom_sf(aes(fill = fittedIV)) +
  facet_wrap(~Year, dir = 'h') +
  ggtitle("Interaction E3CI") + theme_bw() + 
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()
  ) +
  scale_fill_gradient2(
    midpoint = -0, low = "blue", mid = "white", high = "red"
  )
```


#### Model Comparison
For the first three spatio-temporal models we cannot expect good results, as we have seen with the spatial models and correlations that the time-space-covariance relationship is not so simple.

- Basic Spatio-Temporal Models:
```{r}
# Real90_00
Fitted90_00
# Fitted90_00_2
```


- More complex spatio-temporal models: interactions
```{r}
Real90_00
Fitted90_00_I
# Fitted90_00_II
# Fitted90_00_III
Fitted90_00_IV
```

Combine these plots

```{r}
comb_4 <- plot_grid(Real90_00, Fitted90_00_I,ncol = 2) %>% print()
# comb_paramtrend_real <- plot_grid(Real90_00,Fitted90_00,ncol = 2) %>% print()
# comb_int4_real <- plot_grid(Real90_00,Fitted90_00_IV,ncol = 2) %>% print()
```



```{r}
results9099 <- data.frame("E3CI_real" = data90_99$E3CI_diff,"IID" = mod90_99$summary.fitted.values[,"mean"], "rw1" = mod90_99_2$summary.fitted.values[,"mean"], "Interaction_I" = mod90_99_I$summary.fitted.values[,"mean"],"Interaction_II" = mod90_99_II$summary.fitted.values[,"mean"],"Interaction_III" = mod90_99_III$summary.fitted.values[,"mean"],"Interaction_IV" = mod90_99_IV$summary.fitted.values[,"mean"])

mse <- function(real,fitted){
  return(mean((real - fitted)^2))
}
results_df9099 <- cbind(data90_99, results9099)
```

Different available criteria for model selection:

- `MSE` (on train partition)
```{r}
MSE_90_99 <- c(mse(results_df9099$E3CI_diff, fitted = results9099$E3CI_real),
               mse(results_df9099$E3CI_diff, fitted = results9099$IID), 
               mse(results_df9099$E3CI_diff, fitted = results9099$rw1), 
               mse(results_df9099$E3CI_diff, fitted = results9099$Interaction_I),
               mse(results_df9099$E3CI_diff, fitted = results9099$Interaction_II),
               mse(results_df9099$E3CI_diff, fitted = results9099$Interaction_III),
               mse(results_df9099$E3CI_diff, fitted = results9099$Interaction_IV))
#MSE_90_99
```


- `cpo` can be computed in the following way:
```{r}
cpo <- function(inla_model){
  return(-sum(log(inla_model$cpo$cpo)))
}

cpo_iid_time <- cpo(mod90_99)
cpo_rw1 <- cpo(mod90_99_2)
cpo_interactionI <- cpo(mod90_99_I)
cpo_interactionII <- cpo(mod90_99_II)
cpo_interactionIII <- cpo(mod90_99_III)
cpo_interactionIV <- cpo(mod90_99_IV)


CPOs_9099 <- c(cpo_iid_time,cpo_rw1,cpo_interactionI,cpo_interactionII,
               cpo_interactionIII,cpo_interactionIV)
```

- `marginal likelihood`: **Hay dos valores, no se cual es el que pone la teoria**
```{r}
MLIKs9099 <- c(mod90_99$mlik[[1,1]],mod90_99_2$mlik[[1,1]],
               mod90_99_I$mlik[[1,1]],
               mod90_99_II$mlik[[1,1]],
               mod90_99_III$mlik[[1,1]],
               mod90_99_IV$mlik[[1,1]])
```

- `DIC`
```{r}
DICs9099 <- c(mod90_99$dic$dic,mod90_99_2$dic$dic,
              mod90_99_I$dic$dic,
              mod90_99_II$dic$dic,
              mod90_99_III$dic$dic,
              mod90_99_IV$dic$dic)
```

- `WAIC` 
```{r}
WAICs9099 <- c(mod90_99$waic$waic,mod90_99_2$waic$waic,
               mod90_99_I$waic$waic,
               mod90_99_II$waic$waic,
               mod90_99_III$waic$waic,
               mod90_99_IV$waic$waic)
```


```{r}
percentage_diff_sign <- function(vector1, vector2) {
  # Check that the vectors have the same length
  if (length(vector1) != length(vector2)) {
    stop("The vectors must have the same length.")
  }
  
  # Calculate the number of elements with differing signs
  diff_signs <- sum(sign(vector1) != sign(vector2))
  
  # Calculate the percentage
  percentage <- (diff_signs / length(vector1)) * 100
  
  return(percentage)
}
signs90_99 <- c(percentage_diff_sign(results_df9099$E3CI_diff,results_df9099$fitted),
                percentage_diff_sign(results_df9099$E3CI_diff,results_df9099$fitted2),
                percentage_diff_sign(results_df9099$E3CI_diff,results_df9099$fittedI),
                percentage_diff_sign(results_df9099$E3CI_diff,results_df9099$fittedII),
                percentage_diff_sign(results_df9099$E3CI_diff,results_df9099$fittedIII),
                percentage_diff_sign(results_df9099$E3CI_diff,results_df9099$fittedIV))
```

```{r}
errors_df9099_t <- data.frame("Likelihood" = rep("gaussian",6),
                              "Model" = c("bym + IID", "bym + rw1", "Interaction I",
                                        "Interaction II","Interaction III","Interaction IV"), 
               "WAIC" = WAICs9099,
               "MLIK" = MLIKs9099,
               "CPO" = CPOs_9099,
               "DIC" = DICs9099,
               "MSE" = MSE_90_99[2:length(MSE_90_99)],
               "SignError" = signs90_99)

library(kableExtra)
kable(errors_df9099_t)
#xtable::xtable(errors_df9099_t, digits = 3)
```

##### Gaussian vs t-student

```{r}
errors_df9099 <- rbind(errors_df9099_n,errors_df9099_t)
kable(errors_df9099)
```


#### Predictions

First, we save the current `data90_99` which contains the fitted values of the above models
```{r}
data90_99_saved <- data90_99
```

To predict with INLA we only have to include the observations we want to predict **with a missing value in the response**. 

Then, let us reconsider `data_9099` but now we are going to include the year 2000:
```{r}
years9099 <- c(1982:2022)
data90_99 <- NULL
for(i in years9099){
  data90_99 <- rbind(data90_99,select_year_diff2(year = i))
}
data90_99_bench <- data90_99
data90_99[which(data90_99$Year %in% c(2022:2022)),"E3CI_diff"] = NA
```

Now, we refit the best model obtained above: *bym + interaction*.

```{r}
# Create indexes as before
#data90_99$idarea <- as.numeric(as.factor(data90_99$color_code))
data90_99$idarea1 <- data90_99$idarea
#data90_99$idtime <- 1 + data90_99$Year - min(data90_99$Year)
data90_99$idtime1 <- data90_99$idtime
data90_99$idareatime <- interaction(data90_99$idarea,data90_99$idtime)
data90_99$idtime_int <- data90_99$idtime
data90_99$idarea_int <- data90_99$idarea

formula_9090_I <- E3CI_diff ~ 1 + GDP_diff + CO2_diff + FA_diff + 
  f(idarea, model = "bym", graph = g) + 
  f(idtime, model = "rw1") + f(idtime1, model = "iid") +
  f(idareatime, model = "iid")

# Fit the model
mod90_99_predI <- inla(formula = formula_9090_I, 
                 family = "gaussian",
                 data = data90_99,
                 control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE),
                 control.predictor = list(compute = TRUE))
```

Now, Inla has used the predictive distribution to compute the values for the year 2000:

```{r}
rows2000 <- which(data90_99$Year == 2000)
rows2000_bench <- which(data90_99_bench$Year == 2000)
pred_2000 <- mod90_99_predI$summary.fitted.values[rows2000,"mean"]

#preds_mod9099 <- data.frame("Real_2000" = data90_99_bench[rows2000_bench,"E3CI"])
data90_99_bench$fittedI <- mod90_99_predI$summary.fitted.values[,"mean"]
data90_99_bench <- data90_99_bench %>%  relocate(fittedI, .after = E3CI)
```

**INLA does what he say?** 
```{r}
rows95_1 <- which(data90_99_saved$Year == 1995)
rows95_2 <- which(data90_99_bench$Year == 1995)
unique(rows95_2 == rows95_1) # True

# Same fit?
unique(data90_99_saved$fittedI[rows95_1] == data90_99_bench$fittedI[rows95_2])

```



Then, we can check the MSE error of the predicted values:

```{r}
# To check whether the order is correct or not 
unique(rows2000_bench == rows2000)
unique(data90_99$color_code[rows2000] == data90_99_bench$color_code[rows2000_bench])


df_pred90_99 <- data.frame("Real" = data90_99_bench$E3CI_diff[rows2000_bench],
                      "fittedI" = data90_99_bench$fittedI[rows2000_bench],
                      "Code1" = data90_99_bench$color_code[rows2000_bench],
                      "Code2" = data90_99$color_code[rows2000_bench])

#View(df_pred90_99)

MSE_pred2000 <- mse(real = data90_99_bench$E3CI_diff[rows2000_bench],
                    fitted = data90_99_bench$fittedI[rows2000_bench]) %>% print()
```

It could be also of interest to check the number of missclassified signs:
```{r}
percentage_diff_sign <- function(vector1, vector2) {
  # Check that the vectors have the same length
  if (length(vector1) != length(vector2)) {
    stop("The vectors must have the same length.")
  }
  
  # Calculate the number of elements with differing signs
  diff_signs <- sum(sign(vector1) != sign(vector2))
  
  # Calculate the percentage
  percentage <- (diff_signs / length(vector1)) * 100
  
  return(percentage)
}
```


```{r}
map2000_real <- ggplot(data = data90_99_bench[rows2000_bench,]) +
  geom_sf(aes(fill = E3CI_diff)) + 
  labs(fill = "Real ERA5 - 2000") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red")

map2000_predI <- ggplot(data = data90_99_bench[rows2000_bench,]) +
  geom_sf(aes(fill = fittedI)) + 
  labs(fill = "Bym+Interaction ERA5 - 2000") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red")

map2000_real
map2000_predI
Mse_pred2000 <- mse(data90_99_bench$E3CI_diff[rows2000_bench],data90_99_bench$fittedI[rows2000_bench])

signerror2000<- percentage_diff_sign(data90_99_bench$E3CI_diff[rows2000_bench],data90_99_bench$fittedI[rows2000_bench])
```

```{r}
rows2001_bench <- which(data90_99_bench$Year == 2001)
map2001_real <- ggplot(data = data90_99_bench[rows2001_bench,]) +
  geom_sf(aes(fill = E3CI_diff)) + 
  labs(fill = "Real ERA5 - 2001") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red")

map2001_predI <- ggplot(data = data90_99_bench[rows2001_bench,]) +
  geom_sf(aes(fill = fittedI)) + 
  labs(fill = "Bym+Interaction ERA5 - 2001") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red")

map2001_real
map2001_predI
Mse_pred2001 <- mse(data90_99_bench$E3CI_diff[rows2001_bench],data90_99_bench$fittedI[rows2001_bench]) %>% print()

signerror2001 <- percentage_diff_sign(data90_99_bench$E3CI_diff[rows2001_bench],data90_99_bench$fittedI[rows2001_bench]) %>% print()
```


```{r}
rows2002_bench <- which(data90_99_bench$Year == 2002)
map2002_real <- ggplot(data = data90_99_bench[rows2002_bench,]) +
  geom_sf(aes(fill = E3CI_diff)) + 
  labs(fill = "Real ERA5 - 2002") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red")

map2002_predI <- ggplot(data = data90_99_bench[rows2002_bench,]) +
  geom_sf(aes(fill = fittedI)) + 
  labs(fill = "Bym+Interaction ERA5 - 2002") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red")

map2002_real
map2002_predI

Mse_pred2002 <- mse(data90_99_bench$E3CI_diff[rows2002_bench],data90_99_bench$fittedI[rows2002_bench])

signerror2002 <- percentage_diff_sign(data90_99_bench$E3CI_diff[rows2002_bench],data90_99_bench$fittedI[rows2002_bench]) %>% print()
```

```{r}
rows2003_bench <- which(data90_99_bench$Year == 2003)
map2003_real <- ggplot(data = data90_99_bench[rows2003_bench,]) +
  geom_sf(aes(fill = E3CI_diff)) + 
  labs(fill = "Real ERA5 - 2003") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red")

map2003_predI <- ggplot(data = data90_99_bench[rows2003_bench,]) +
  geom_sf(aes(fill = fittedI)) + 
  labs(fill = "Bym+Interaction ERA5 - 2003") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red")

map2003_real
map2003_predI
Mse_pred2003 <- mse(data90_99_bench$E3CI_diff[rows2003_bench],data90_99_bench$fittedI[rows2003_bench]) %>% print()
signerror2003 <- percentage_diff_sign(data90_99_bench$E3CI_diff[rows2003_bench],data90_99_bench$fittedI[rows2003_bench]) %>% print()
```

```{r}
rows2004_bench <- which(data90_99_bench$Year == 2004)
map2004_real <- ggplot(data = data90_99_bench[rows2004_bench,]) +
  geom_sf(aes(fill = E3CI_diff)) + 
  labs(fill = "Real ERA5 - 2004") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red")

map2004_predI <- ggplot(data = data90_99_bench[rows2004_bench,]) +
  geom_sf(aes(fill = fittedI)) + 
  labs(fill = "Bym+Interaction ERA5 - 2004") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red")

map2004_real
map2004_predI

Mse_pred2004 <- mse(data90_99_bench$E3CI_diff[rows2004_bench],data90_99_bench$fittedI[rows2004_bench]) %>% print()

signerror2004 <- percentage_diff_sign(data90_99_bench$E3CI_diff[rows2004_bench],data90_99_bench$fittedI[rows2004_bench]) %>% print()
```


```{r}
rows2005_bench <- which(data90_99_bench$Year == 2005)
map2005_real <- ggplot(data = data90_99_bench[rows2005_bench,]) +
  geom_sf(aes(fill = E3CI_diff)) + 
  labs(fill = "Real ERA5 - 2005") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red")

map2005_predI <- ggplot(data = data90_99_bench[rows2005_bench,]) +
  geom_sf(aes(fill = fittedI)) + 
  labs(fill = "Bym+Interaction ERA5 - 2005") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red")

map2005_real
map2005_predI
Mse_pred2005 <- mse(data90_99_bench$E3CI_diff[rows2005_bench],data90_99_bench$fittedI[rows2005_bench]) %>% print()

signerror2005 <- percentage_diff_sign(data90_99_bench$E3CI_diff[rows2005_bench],data90_99_bench$fittedI[rows2005_bench]) %>% print()
```

All joint:

```{r}
filtered_data <- data90_99_bench %>%
  filter(Year %in% c(2001, 2002, 2003, 2004, 2005))

combined_df <- filtered_data %>%
  select(geometry, Year, E3CI_diff, fittedI) %>%
  pivot_longer(cols = c(E3CI_diff, fittedI), 
               names_to = "Variable", 
               values_to = "Value") %>%
  unite("Year_Variable", Year, Variable, sep = "_")

combined_map <- ggplot(combined_df) +
  geom_sf(aes(fill = Value)) + 
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red") +
  facet_wrap(~ Year_Variable, ncol = 5) +
  labs(fill = "Value") +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        strip.background = element_blank(),
        strip.text = element_text(face = "bold", size = 12),
        legend.position = "right") 

print(combined_map)
```

In a table
```{r}
preds_df9099 <- data.frame(Year = 2000:2005,
                           SignError = c(signerror2000,signerror2001,
                                          signerror2002,signerror2003,
                                          signerror2004,signerror2005),
                           MSE = c(Mse_pred2000,Mse_pred2001,
                                    Mse_pred2002,Mse_pred2003,
                                    Mse_pred2004,Mse_pred2005))

# transformed_df <- data.frame(
#   Metric = c("SignError","SignError","MSE","MSE"),
#   Model = c("Spatial", "Spatio-Temp","Spatial", "Spatio-Temp"),
#   `2000` = c(signerror2000_00, signerror2000, MSE_2000_00, Mse_pred2000),
#   `2001` = c(signerror2001_00, signerror2001, MSE_2001_00, Mse_pred2001),
#   `2002` = c(signerror2002_00, signerror2002, MSE_2002_00, Mse_pred2002),
#   `2003` = c(signerror2003_00, signerror2003, MSE_2003_00, Mse_pred2003),
#   `2004` = c(signerror2004_00, signerror2004, MSE_2004_00, Mse_pred2004),
#   `2005` = c(signerror2005_00, signerror2005, MSE_2005_00, Mse_pred2005)
# )

transformed_df <- data.frame(
  Metric = c("SignError","SignError","MSE","MSE"),
  Model = c("Spatial", "Spatio-Temp","Spatial", "Spatio-Temp"),
  `2000` = c(NA, signerror2000, NA, Mse_pred2000),
  `2001` = c(NA, signerror2001, NA, Mse_pred2001),
  `2002` = c(NA, signerror2002, NA, Mse_pred2002),
  `2003` = c(NA, signerror2003, NA, Mse_pred2003),
  `2004` = c(NA, signerror2004, NA, Mse_pred2004),
  `2005` = c(NA, signerror2005, NA, Mse_pred2005)
)

kable(transformed_df)
xtable(transformed_df, digits = 3)
```

##### All the sign errors
```{r}
compute_summary <- function(years){
  sign_errors <- NULL
  mses <- NULL
  for(i in years){
    cond <- which(data90_99_bench$Year == i)
    sign_errors <- c(sign_errors,
                     percentage_diff_sign(data90_99_bench$E3CI_diff[cond],
                                          data90_99_bench$fittedI[cond]))
    mses <- c(mses, 
              mse(data90_99_bench$E3CI_diff[cond],
                  data90_99_bench$fittedI[cond]))
  }
  df <- data.frame(Year = years,
                   SignError = sign_errors,
                   MSE = mses)
  return(df)
}
```

```{r}
del82_05_22 <- compute_summary(years9099)
kable(compute_summary(years9099))
xtable(del82_05_22, digits = 7)

```


Problematic years 2014, 2015, 2018 and 2020.

```{r}
outliers <- which((data90_99_bench$E3CI_diff >= 4) | (data90_99_bench$E3CI_diff <= -4))
outliers_GDP <- which((data90_99_bench$GDP_diff >= 4) | (data90_99_bench$GDP_diff <= -4))
#View(data90_99_bench[outliers,c("name","E3CI_diff","fittedI","Year","GDP_diff","CO2_diff","FA_diff")])
```

##### 2020, 2021 and 2022

```{r}
rows2020_bench_e <- which(data90_99_bench$Year == 2020)
map2020_real_e <- ggplot(data = data90_99_bench[rows2020_bench_e,]) +
  geom_sf(aes(fill = E3CI_diff)) + 
  labs(fill = "Real ERA5 - 2021") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red")

map2020_predI_e <- ggplot(data = data90_99_bench[rows2020_bench_e,]) +
  geom_sf(aes(fill = fittedI)) + 
  labs(fill = "Bym+Interaction ERA5 - 2022") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red")

map2020_real_e
map2020_predI_e
Mse_pred2020 <- mse(data90_99_bench$E3CI_diff[rows2020_bench_e],data90_99_bench$fittedI[rows2020_bench_e]) %>% print()

signerror2020 <- percentage_diff_sign(data90_99_bench$E3CI_diff[rows2020_bench_e],data90_99_bench$fittedI[rows2020_bench_e]) %>% print()
```

```{r}
rows2021_bench_e <- which(data90_99_bench$Year == 2021)
map2021_real_e <- ggplot(data = data90_99_bench[rows2021_bench_e,]) +
  geom_sf(aes(fill = E3CI_diff)) + 
  labs(fill = "Real ERA5 - 2021") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red")

map2021_predI_e <- ggplot(data = data90_99_bench[rows2021_bench_e,]) +
  geom_sf(aes(fill = fittedI)) + 
  labs(fill = "Bym+Interaction ERA5 - 2022") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red")

map2021_real_e
map2021_predI_e
Mse_pred2021 <- mse(data90_99_bench$E3CI_diff[rows2021_bench_e],data90_99_bench$fittedI[rows2021_bench_e]) %>% print()

signerror2021 <- percentage_diff_sign(data90_99_bench$E3CI_diff[rows2021_bench_e],data90_99_bench$fittedI[rows2021_bench_e]) %>% print()
```

```{r}
rows2022_bench_e_t <- which(data90_99_bench$Year == 2022)

map2022_real_e_t <- ggplot(data = data90_99_bench[rows2022_bench_e_t,]) +
  geom_sf(aes(fill = E3CI_diff)) + 
  labs(fill = "Real ERA5 - 2022") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red") + 
  theme(legend.position = "left",
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA)
  )

map2022_predI_e_t <- ggplot(data = data90_99_bench[rows2022_bench_e_t,]) +
  geom_sf(aes(fill = fittedI)) + 
  labs(fill = "Interaction I - 2022") +
  scale_fill_gradient2(
    midpoint = 0, low = "blue", mid = "white", high = "red") + 
  theme(legend.position = "top",
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA)
  )

map2022_real_e_t
map2022_predI_e_t
Mse_pred2022_t <- mse(data90_99_bench$E3CI_diff[rows2022_bench_e_t],data90_99_bench$fittedI[rows2022_bench_e_t]) %>% print()

signerror2022_t <- percentage_diff_sign(data90_99_bench$E3CI_diff[rows2022_bench_e_t],data90_99_bench$fittedI[rows2022_bench_e_t]) %>% print()
```


```{r}
comb2022 <- (map2022_real_e_t + map2022_predI_e_t + map2022_predI_e) 
print(comb2022)
```


##### Posterior Information

```{r}
summary(mod90_99_I)
mod90_99_I$summary.fixed[,1:6]
mod90_99_I$summary.hyperpar

xtable(mod90_99_I$summary.fixed[,1:6], digits = 4)
xtable(mod90_99_I$summary.hyperpar)
```

#### Model Hypothesis
Let us check some standard model hypotheses

##### Gaussian Likelihood

```{r}
qqnorm(data90_99$E3CI_diff)
qqline(data90_99$E3CI_diff,col="red")
shapiro.test(data90_99$E3CI_diff)
```


##### Residuals Normality

```{r}
resid_9099 <- data90_99_saved$fittedI - data90_99_saved$E3CI_diff

qqnorm(resid_9099)
qqline(resid_9099,col="red")
shapiro.test(resid_9099)
```
We observe that, there normality is not specially fulfilled in the lower tails, but the p-value is greater than a 0.05 significance level, so that we can not discard normality.

##### Homocedasticity

```{r}
results_df9099$residuals <- resid_9099
#Plot standardized residuals vs fitted values
homoscedasticity_plot <- ggplot(results_df9099, aes(x = fittedI, y = residuals)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "blue") +
  labs(title = "Standardized Residuals vs Fitted Values",
       x = "Fitted Values",
       y = "Standardized Residuals") +
  theme_minimal() +
  theme(panel.grid = element_blank(), axis.text = element_blank(), axis.ticks = element_blank())

print(homoscedasticity_plot)
```



##### Linearity

```{r}
results_df9099$residuals<- resid_9099
results_df9099 <- results_df9099 %>%
  mutate(standardized_residuals = residuals / sd(residuals))

plot1 <- ggplot(results_df9099, aes(x = GDP_diff, y = standardized_residuals)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red") +
  labs(title = "Residuals vs GDP_diff", x = "GDP_diff", y = "Residuals") +
  theme_minimal()

plot2 <- ggplot(results_df9099, aes(x = CO2_diff, y = standardized_residuals)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red") +
  labs(title = "Residuals vs CO2_diff", x = "CO2_diff", y = "Residuals") +
  theme_minimal()

plot3 <- ggplot(results_df9099, aes(x = FA_diff, y = standardized_residuals)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red") +
  labs(title = "Residuals vs FA_diff", x = "FA_diff", y = "Residuals") +
  theme_minimal()

# Combine the plots into a single plot
# combined_plot <- plot_grid(plot1, plot2, plot3, ncol = 2)
combined_plot <- plot_grid(plot1, plot2, plot3, homoscedasticity_plot, ncol = 2)

# Display the combined plot
print(combined_plot)
```

The linearity is not fulfilled at all. But there is no a clear patterns to think that the values can not be randomly distributed around the line.


##### Independence

```{r}
library(lmtest)

# Durbin-Watson
dw_test <- dwtest(resid_9099 ~ mod90_99_I$summary.fitted.values[,"mean"])
print(dw_test)
```


##### Outliers/Gaussian Likelihood
The models that INLA fits assume a lot of structure LGM+GMRF. However, this model specification allows for different likelihoods. We have specified `family = gaussian` in each model. This allows us to check whether the observed response values are likely to follow a normal distribution. We proceed with a *shapiro test*. The null hypothesis is then that the sample is normally distributed.

```{r}
test2000 <- shapiro.test(data90_99$E3CI_diff)
test2000$p.value
```

The p-value of `r test2000$p.value` says that there is no enough evidence to reject that our sample comes from a Normal population.

```{r}
qqnorm(data90_99$E3CI_diff)
qqline(data90_99$E3CI_diff, col = "red")
```









